---
title: |
        **NEW**  \
        Figure 3 & Associated Supplemental Figures
subtitle: "Model Validation and Feature Annotation"
format: 
  html:
    fig-format: svg
    filters:
      - inline-svg
editor: source
author: "C.P. Shannon"
execute:
  echo:    false
  warning: false
  message: false
self-contained: true
self-contained-math: true
---

```{r setup}
#| include: false

set.seed(71829)

library(tidyverse)
ggplot2::theme_set(
  cowplot::theme_cowplot() + 
    ggplot2::theme(
      strip.background = ggplot2::element_blank(),
      axis.title    = element_text(size = 10),
      plot.title    = element_text(face = 'bold', size = 11),
      plot.subtitle = element_text(size = 11)
    )
)

# potential sample swaps in png
id_sex <- c('6F1A', '5A8C', '9J3D', '2G7B')
id_dol <- c('1D8A', '7D5D', '2D1D', '9D4K', '4M5A', '3G8H')

my_pal <- c("#E64B35FF", "#4DBBD5FF", "#00A087FF", "#3C5488FF")
proj3cols <-  c('#3C5488FF', '#6F286AFF', '#E64B35FF', 'darkgrey')

# helper - ggplot
StatAreaUnderDensity <- ggproto('StatAreaUnderDensity', Stat, required_aes = 'x', compute_group = function(data, scales, xlim = NULL, n = 50) {
    fun <- approxfun(density(data$x))
    StatFunction$compute_group(data, scales, fun = fun, xlim = xlim, n = n)
  })

# helper - ggplot
stat_aud <- function(mapping = NULL, data = NULL, geom = "area", position = "identity", na.rm = FALSE, show.legend = NA, inherit.aes = TRUE, n = 50, xlim = NULL, ...) {
  layer(
    stat = StatAreaUnderDensity, data = data, mapping = mapping, geom = geom, 
    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
    params = list(xlim = xlim, n = n, ...)
  )
}

# helper - scatterplot 
my_scatter <- function(data, scale = T, group = NULL, nboot = NULL, col = NULL, site = 'NA', ...) {

  # rescale?
  if(scale)
    data <- data %>% mutate(across(c(actual, prediction), scales::rescale))

  # bootstrap to obtain cor, 95% CI and empirical p-value for hypothesis that cor>0
  if(!is.null(nboot)) {
    stats <- data %>%
      rsample::bootstraps(times = nboot) %>%
      mutate(perf = map_dbl(splits, ~ cor(as_tibble(.x)$prediction, as_tibble(.x)$actual, method = 'spearman'))) %>%
      summarize(mean = median(perf, na.rm = T), lower = quantile(perf, 0.05, na.rm = T), upper = quantile(perf, 0.95, na.rm = T), pval = ecdf(perf)(0))
  } else {
    stats <- tibble(mean = cor(data$prediction, data$actual, method = 'spearman'))
  }

  # plot
  gg <- data %>% ggplot(aes(actual, prediction))

  # colour?
  if(!is.null(col))
    gg <- gg + geom_point(size = 2, pch = 21, col = col)
  else
    gg <- gg + geom_point(size = 2, pch = 21)

  # cvfolds?
  if(!is.null(group))
    gg <- gg + stat_smooth(method = MASS::rlm, alpha = 0.1, size = 0.2, colour = 'black', fullrange = T, linetype = 'dashed', se = F, aes(group = .data[[group]]))
  
  # overall trend line
  gg <- gg + stat_smooth(method = MASS::rlm, alpha = 0.1, size = 0.5, colour = 'black', fullrange = T)

  if(scale) {
    gg <- gg +
      coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
      annotate(geom = 'text', label = sprintf('n=%d', nrow(data)), x = 1, y = 0, hjust = 1, vjust = 0)
  } else {
    gg <- gg + coord_cartesian(xlim = c(min(c(data$actual, data$prediction)), NA), ylim = c(min(c(data$actual, data$prediction)), NA))
  }

  gg + labs(x = 'actual', y = 'predicted')

  if(is.null(group) & is.null(nboot))
    subtitle <- with(stats, sprintf('train\nrho=%.2f', mean))
  if(!is.null(group))
    subtitle <- with(stats, sprintf('k-fold cv\nrho=%.2f [%.2f-%.2f]', list(...)$cvm, list(...)$cvlo, list(...)$cvup))
  if(!is.null(nboot))
    subtitle <- with(stats, sprintf('test: %s\nrho=%.2f [%.2f-%.2f]', site, mean, lower, upper))
    # subtitle <- with(stats, sprintf('test: %s\nrho=%.2f [%.2f-%.2f]; p=%.2g', site, mean, lower, upper, pval))

  gg <- gg + labs(subtitle = subtitle) + theme(axis.text = element_text(size = 10))

  return(gg)
}

# helper - bootstrap null model
my_boot_mae <- function(data, nboot = 2000, ...) {
  # filter out potential sample-swaps (predicted sex-mismatch)
  data <- data %>% filter(!vid %in% id_sex)
  
  err_model = yardstick::mae(data, truth = actual, estimate = prediction) %>% pull(.estimate)
  boot <- rsample::bootstraps(data, nboot)
  boot <- boot$splits %>%
    map(rsample::analysis) %>%
    map(mutate, null = 0) %>%
    map(summarize, err_null  = yardstick::mae_vec(truth = actual, estimate = null)) %>% 
    map_dbl(pull, err_null)
  
  x_min <- -0.1
  x_max <- 1.3
  col <- ggsci::pal_npg()(2)
  ci_lo <- quantile(boot, na.rm = T, probs = c(0.025))
  
  tibble(err_null = boot) %>% 
    ggplot(aes(err_null)) +
    geom_density() +
    expand_limits(x = c(x_min, x_max)) +
    stat_aud(geom = 'area', fill = col[1], xlim = c(x_min, ci_lo), alpha = 0.2) +
    
    annotate(x = err_model,    y = Inf, geom = 'text', label.r = unit(0, 'pt'), size = 3, hjust =  1.1, vjust = 1.1, label = 'model 🡆') +
    annotate(x = median(boot), y = 3.0, geom = 'text', label.r = unit(0, 'pt'), size = 3, hjust = -0.2, vjust = 1.1, label = '   🡄 null') +
    annotate(geom = 'text', label = sprintf('n=%d', nrow(data)), x = Inf, y = 0, hjust = 1.25, vjust = -0.25) +

    geom_vline(xintercept = err_model, colour = 'black', linetype = 'solid') +
    geom_vline(xintercept = ci_lo, colour = col[1], linetype = 'dotted') +
    labs(x = 'mean absolute error', y = 'density', subtitle = 'test: PNG') +
    theme(axis.text = element_text(size = 10))
}

results <- list()

```

```{r import}
#| echo: false
#| warning: false
#| message: false

con <- tribble(
  ~site,    ~data,
  'gambia', readr::read_rds('data/processed/gam/import_responses.rds'),
  'png',    readr::read_rds('data/processed/png/import_responses.rds')
)

df_titers <- con %>%
  unnest(data) %>%
  rename(subject_id = `Subject ID`) %>% 
  mutate(status = factor(V1 < 2.5, levels = c(F, T), labels = c('HBsAb MA+', 'HBsAb MA-'))) %>% 
  select(-MAT) %>% 
  na.omit() %>% 
  gather(time, titer, V1, V3, V4) %>% 
  mutate(time = factor(time, levels = c('V1', 'V3', 'V4'), labels = c('0', '30', '128')))

```

```{r load-models}
#| echo: false
#| warning: false
#| message: false

# select models
df <- readr::read_rds('results/other_outputs/trained_models.rds')

# load cv performance summary table
tmp <- readr::read_rds('results/other_outputs/diablo_variablecv_top20p_results_summarized_rho.rds') %>% group_by(across(paired:response)) %>% slice_max(order_by = cvm, n = 1) %>% ungroup()
df <- inner_join(tmp, df)

# bring in training and cv details - abridged only includes results where: dol == 0, daygrp == 'DOL7', grp == 'ALL', status != 'ALL', nkeep == 2000
tmp <- readr::read_rds('results/other_outputs/diablo_variablecv_top20p_results_abridged.rds') %>% select(paired:response, ncomp:cv_type, results) %>% inner_join(df, .)
df <- inner_join(tmp, df)

# re-order
df <- df %>% arrange(status, response)

# add colour - using ggsci::pal_npg()
df$colour <- df %>% select(status, response) %>% unite(col, status, response) %>% pull(col) %>% factor(labels = my_pal) %>% as.character()

# container for plots
figs <- df %>% select(paired:response)

```

```{r prep-titers}
#| echo: false
#| warning: false
#| message: false

titers <- list()

# GAM
a <- readr::read_rds('data/processed/gam/import_metadata.rds')  %>% 
  select(uid = `Unique Identifier`, vid = `Visit Identifier`, sex = Sex, grp = VaccineGrp, dol = DOL)

b <- readr::read_rds('data/processed/gam/import_responses.rds') %>% 
  mutate('log(V3/V1)' = log(V3)-log(V1), 'log(V4/V1)' = log(V4)-log(V1), V1 = log(V1)) %>% 
  select(uid = `Subject ID`, V1, starts_with('log'))

titers$gam <- full_join(a, b)
rm(a, b)

# PNG
a <- readr::read_rds('data/processed/png/import_metadata.rds')  %>% 
  select(uid = Unique_Identifier, vid = VID, sex = Sex, grp = VaccineGroup, dol = DOL)

b <- readr::read_rds('data/processed/png/import_responses.rds') %>% 
  mutate('log(V3/V1)' = log(V3)-log(V1), 'log(V4/V1)' = log(V4)-log(V1), V1 = log(V1)) %>% 
  select(uid = `Subject ID`, V1, starts_with('log'))


titers$png <- full_join(a, b)
rm(a, b)

titers <- bind_rows(titers, .id = 'site') %>% 
  mutate(status = factor(exp(V1) <= 2.5, levels = c(F, T), labels = c('NON-NAIVE', 'NAIVE')) %>% as.character()) %>% 
  gather(response, actual, starts_with('log'))

```

```{r perf-train}
#| echo: false
#| warning: false

df$training <- df %>% 
  pmap(function(colour, comp, results, ...) {
    results %>% 
      filter(type == 'training', comp == !!comp) %>% 
      unnest(assess) %>% 
      my_scatter(col = colour)
  })

figs <- df %>% select(paired:response, training) %>% inner_join(figs)

```

```{r perf-cv}
#| echo: false
#| warning: false

df$cv <- df %>%
  pmap(function(cvm, cvlo, cvup, colour, comp, results, ...) {
    
    gg <- results %>% 
      filter(type == 'cv', comp == !!comp) %>% 
      unnest(assess)
    
    gg_examplar <- gg %>%
      group_by(id) %>%
      summarise(cvm = cor(actual, prediction, method = 'spearman')) %>%
      mutate(delta = abs(cvm - !!cvm)) %>% slice_min(delta, n = 1) %>%
      inner_join(gg)
    
    gg_examplar %>% my_scatter(scale = T, group = 'id2', col = colour, cvm = cvm, cvlo = cvlo, cvup = cvup)

  })

figs <- df %>% select(paired:response, cv) %>% inner_join(figs)

```

```{r perf-discord}
#| echo: false
#| warning: false

# predict on everyone at the appropriate DOL
# subset to corresponding HBsAb status based on joining with metadata data...
discord <- readRDS('data/processed/cross_validation/data_master_set_top20p.rds')
discord <- discord %>%
  filter(!paired) %>% 
  mutate(data = set_names(data, block)) %>% 
  nest(x = c(block, data)) %>% 
  mutate(x = map(x, pull, data))

discord <- df %>% 
  select(-c(x, y, outcome, results)) %>% 
  inner_join(discord)
  
discord$pred <- discord %>% 
  pmap(function(diablo, comp, x, ...) {
    
    # create subset and pad data according to model
    x <- tibble(vars = map(diablo$X[names(x)], colnames), x = x) %>%
      pmap(function(vars, x, ...) {
        missing <- vars[!vars %in% colnames(x)]
        mat <- matrix(NA, nrow = nrow(x), ncol = length(missing), dimnames = list(rownames(x), missing))
        mat <- cbind(x, mat)
        mat[ , vars]
      })
    
    # only use complete omic observations across selected blocks    
    i <- x %>% map(rownames) %>% reduce(intersect)
    x <- x %>% map(~ .[i, ])
    
    depth <-  comp %>% gsub('comp', '', .) %>% as.numeric()
    pred <- mixOmics:::predict.block.spls(object = diablo, newdata = x)$AveragedPredict[ , 1, depth]
    
    tibble(vid = names(pred), prediction = pred)
  })

tmp <- titers %>% 
  select(site, uid, vid, dol, status, response, actual) %>% 
  group_by(uid) %>% 
  mutate(daygrp = sprintf('DOL%d', max(dol))) %>% 
  ungroup()

discord <- discord %>% 
  select(paired:response, pred) %>% 
  unnest(pred) %>% 
  left_join(tmp, by = c('dol', 'response', 'vid'), suffix = c('_model', '_test')) %>%
  filter(
    daygrp_test == daygrp_model,
    status_test != status_model
  ) %>% 
  na.omit() %>%
  select(status_model, daygrp_model, daygrp_test, response, uid, vid, prediction, actual) %>% 
  nest(data = c(uid, vid, daygrp_model, daygrp_test, prediction, actual))

my_colours <- ggsci::pal_npg()(4)
names(my_colours) <- c('NAIVE log(V3/V1)', 'NAIVE log(V4/V1)', 'NON-NAIVE log(V3/V1)', 'NON-NAIVE log(V4/V1)')

discord$discord <- discord %>%
  pmap(function(data, status_model, response, ... ) {
    data %>% 
      mutate(across(c(actual, prediction), scales::rescale)) %>% 
      ggplot(aes(actual, prediction)) +
      geom_point(colour = my_colours[paste(status_model, response)], pch = 21, size = 2, show.legend = F) +
      geom_point(size = 0.5) +
      stat_smooth(method = MASS::rlm, alpha = 0.1, size = 0.5, colour = 'black', fullrange = T) +
      ggpubr::stat_cor(output.type = 'text', method = 'spearman', aes(label = sprintf('rho=%.2f', after_stat(r))), show.legend = F) +
      ggsci::scale_color_npg() +
      coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
      theme(legend.position = 'top', axis.text = element_text(size = 10))
  })

figs <- discord %>% 
  select(status = status_model, response, discord) %>% 
  left_join(figs, .)

```

```{r perf-test-gam}
#| echo: false
#| warning: false

# predict on everyone at the appropriate DOL
# subset to corresponding HBsAb status based on joining with metadata data...

# from 01_validation_wrangle_gam.qmd
test <- readr::read_rds('data/processed/process_validation/data_gam_incomplete_omic_profiles.rds')

test <- test %>% 
  mutate(data = set_names(data, block)) %>% 
  nest(x = c(block, data)) %>% 
  mutate(x = map(x, pull, data))

test <- df %>% 
  select(-c(x, y, outcome, results)) %>% 
  inner_join(test)
  
test$pred <- test %>% 
  pmap(function(diablo, comp, x, ...) {
    
    # create subset and pad data according to model
    x <- tibble(vars = map(diablo$X[names(x)], colnames), x = x) %>%
      pmap(function(vars, x, ...) {
        missing <- vars[!vars %in% colnames(x)]
        mat <- matrix(NA, nrow = nrow(x), ncol = length(missing), dimnames = list(rownames(x), missing))
        mat <- cbind(x, mat)
        mat[ , vars]
      })
    
    # only use complete omic observations across selected blocks    
    i <- x %>% map(rownames) %>% reduce(intersect)
    x <- x %>% map(~ .[i, ])
    
    depth <-  comp %>% gsub('comp', '', .) %>% as.numeric()
    pred <- mixOmics:::predict.block.spls(object = diablo, newdata = x)$AveragedPredict[ , 1, depth]
    
    tibble(vid = names(pred), prediction = pred)
  })

# save for later
results$gam <- test

# shuffle
test <- test %>% 
  select(paired:response, pred) %>% 
  unnest(pred) %>% 
  inner_join(select(titers, site, uid, vid, dol, status, response, actual)) %>% 
  na.omit() %>% 
  nest(data = c(uid, vid, prediction, actual))

# add colour - using ggsci::pal_npg()
test$colour <- test %>% select(status, response) %>% unite(col, status, response) %>% pull(col) %>% factor(labels = my_pal) %>% as.character()

# bootstrap to obtain cor, 95% CI and empirical p-value for hypothesis that cor>0
test$test_gam <- test %>% pmap(\(data, colour, ...) my_scatter(data, nboot = 2000, site = 'GAM', col = colour))

figs <- test %>% select(paired:response, test_gam) %>% inner_join(figs)

```

```{r perf-test-png}
#| echo: false
#| warning: false

# predict on everyone at the appropriate DOL
# subset to corresponding HBsAb status based on joining with metadata data...

# from 01_validation_wrangle_png.qmd
test <- readr::read_rds('data/processed/process_validation/data_png_combat_corrected.rds')

# only luminex and transcriptomics
test <- test %>% filter(block == 'luminex_cytokines' | block == 'transcriptomics')

test <- test %>% 
  mutate(data = set_names(data, block)) %>% 
  nest(x = c(meta, block, data)) %>% 
  mutate(x = map(x, pull, data))

test <- df %>% 
  select(-c(x, y, outcome, results)) %>% 
  inner_join(test)

test$pred <- test %>% 
  pmap(function(diablo, comp, x, ...) {
    
    # create subset and pad data according to model
    x <- tibble(vars = map(diablo$X[names(x)], colnames), x = x) %>%
      pmap(function(vars, x, ...) {
        missing <- vars[!vars %in% colnames(x)]
        mat <- matrix(NA, nrow = nrow(x), ncol = length(missing), dimnames = list(rownames(x), missing))
        mat <- cbind(x, mat)
        mat[ , vars]
      })

    # only use complete omic observations across selected blocks    
    i <- x %>% map(rownames) %>% reduce(intersect)
    x <- x %>% map(~ .[i, ])
    
    depth <-  comp %>% gsub('comp', '', .) %>% as.numeric()
    pred <- mixOmics:::predict.block.spls(object = diablo, newdata = x)$AveragedPredict[ , 1, depth]

    tibble(vid = names(pred), prediction = pred)
  })

# save predictions for later
results$png <- test

# shuffle
test <- test %>% 
  select(paired:response, pred) %>% 
  unnest(pred) %>% 
  inner_join(select(titers, site, uid, vid, dol, status, response, actual)) %>% 

  # filter out potential sample-swaps (predicted sex-mismatch)
  filter(!vid %in% id_sex) %>%
  
  na.omit() %>% 
  nest(data = c(uid, vid, prediction, actual))

# add colour - using ggsci::pal_npg()
test$colour <- test %>% select(status, response) %>% unite(col, status, response) %>% pull(col) %>% factor(labels = my_pal) %>% as.character()

# different functions based on response visit
test <- test %>% mutate(fn = ifelse(grepl('V3', response), list(my_boot_mae), list(my_scatter)))

# scatterplots for log(V4/V1), dist. of null model performance for log(V3/V1)
# bootstrap to obtain cor, 95% CI and empirical p-value for hypothesis that cor>0
test$gg <- test %>% pmap(\(fn, colour, data, mean, lower, upper, pval, ...) exec(fn, data, nboot = 2000, site = 'PNG', col = colour))

figs <- test %>% select(paired:response, test_png = gg) %>% inner_join(figs)

```

```{r label-sepsis-png}

# id hospitalized newborns to label them on plot
verify <- aws.s3::s3read_using(readr::read_csv, object = 's3://epichipc-main/Clean_Data/PNG_validation/PNGVAL_SINGLEOMICS_CLINICAL.csv') %>%
# verify <- readr::read_csv('data/raw/png/PNGVAL_SINGLEOMICS_CLINICAL.csv') %>%
  select(uid = Unique_Identifier, hosp = Notes) %>%
  filter(!is.na(hosp))

# ugly hack to fix existing plot...
tmp <- figs$test_png[[4]]
tmp_df <- tmp$data %>% inner_join(verify)
tmp <- tmp + 
  ggrepel::geom_text_repel(data = tmp_df, aes(label = hosp), min.segment.length = 0, nudge_x = 0.15, nudge_y = -0.1, size = 3) +
  geom_point(data = tmp_df, size = 0.5)

# now put it back...
figs$test_png[[4]] <- tmp

# cleanup
rm(tmp, tmp_df)

```

```{r auc}

# classification performance in test:
# 
# here we quantify the ability of DOL128 models to discriminate neonates with
# HBsAb titers after the 3-dose series above or below the mean titer value,
# corresponding to ~1000.

set.seed(997)

# auc from raw
titers <- con$data %>% 
  set_names(c('GAM', 'PNG')) %>% 
  map(select, uid = `Subject ID`, actual = V4) %>% 
  bind_rows(.id = 'site') %>% 
  drop_na()

my_tests <- figs %>% 
  filter(response == 'log(V4/V1)') %>%
  arrange(desc(status)) %>% 
  select(paired, dol, daygrp, status, response, starts_with('test')) %>% 
  pmap(\(test_gam, test_png, status, ...) list('PNG' = test_png$data, 'GAM' = test_gam$data) %>% bind_rows(.id = 'site') %>% mutate(status = factor(status, levels = c('NAIVE', 'NON-NAIVE'), labels = c('MA-', 'MA+')))) %>% 
  bind_rows() %>% select(-actual) %>% inner_join(titers) %>% 
  nest(data = -c(site, status)) %>%
  unite(id, status, site, sep = ' ') %>% mutate(data = set_names(data, id)) %>% 
  pull(data)

# remove sepsis
my_tests <- my_tests %>% map(anti_join, filter(verify, hosp == 'sepsis'))

# add aggregate MA-, MA+, and overall test results
my_tests$`ALL` <- bind_rows(my_tests)
my_tests$`MA+` <- bind_rows(my_tests[1:2])
my_tests$`MA-` <- bind_rows(my_tests[3:4])

lim_png <- my_tests[c(1, 3)] %>% bind_rows() %>% pull(actual) %>% log10() %>% mean()
lim_all <- my_tests[1:4]     %>% bind_rows() %>% pull(actual) %>% log10() %>% mean()

# ability to discriminate the upper quartile of responders from all others
fig_roc <- my_tests %>% 
  imap(~ {

    limit <- lim_all
    if(grepl('MA\\+ PNG', .y))
      limit <- lim_png

    df <- .x %>% mutate(label = log10(actual) >= limit)

    # p-value from Mann-Whitney U test
    w <- wilcox.test(prediction ~ label, data = df, alternative = 'less', exact = F)
    auc <- (mean(rank(df$prediction)[df$label == TRUE]) - (sum(df$label) + 1)/2)/(nrow(df) - sum(df$label))
    
    # roc to plot
    roc <- pROC::roc(label ~ prediction, data = df, direction = '<')

    list(roc = roc, label = sprintf('%s (n=%d)\nAUC=%.2f; p=%.1g', .y, nrow(df), auc, w$p.value))
  })

# now create the 2 plots
# MA-
fig_ma_neg <- fig_roc[c('MA- PNG', 'MA- GAM', 'MA-')]
p1 <- fig_ma_neg %>% 
  map('roc') %>% 
  pROC::ggroc()

p1_lab <- fig_ma_neg %>% map('label')
p1_tit <- p1_lab %>% last() %>% unlist()
p1_lab <- p1_lab %>% head(-1) %>%  enframe() %>% unnest(value)

p1 <- p1 +
  geom_line(aes(alpha = name == 'MA-'), show.legend = F) +
  geom_text(data = p1_lab, aes(label = value), x = 0, y = c(0, 0.15), size = 3, hjust = 1, vjust = -0.1, lineheight = 0.75, show.legend = F) +
  scale_alpha_discrete(range = c(0.3, 1)) +
  scale_colour_manual(values = c('#009E73', '#CC79A7', 'black')) +
  guides(colour = 'none') +
  labs(subtitle = p1_tit) +
  theme(axis.text = element_text(size = 10), axis.title = element_text(size = 10))

# remove default lines to leave modified ones with alpha value
p1$layers <- p1$layers[-1]

# MA+
fig_ma_plu <- fig_roc[c('MA+ PNG', 'MA+ GAM', 'MA+')]
p2 <- fig_ma_plu %>% 
  map('roc') %>% 
  pROC::ggroc()

p2_lab <- fig_ma_plu %>% map('label')
p2_tit <- p2_lab %>% last() %>% unlist()
p2_lab <- p2_lab %>% head(-1) %>%  enframe() %>% unnest(value)

p2 <- p2 +
  geom_line(aes(alpha = name == 'MA+'), show.legend = F) +
  geom_text(data = p2_lab, aes(label = value), x = 0, y = c(0, 0.15), size = 3, hjust = 1, vjust = -0.1, lineheight = 0.75, show.legend = F) +
  scale_alpha_discrete(range = c(0.3, 1)) +
  scale_colour_manual(values = c('#E69F00', '#56B4E9', 'black')) +
  guides(colour = 'none') +
  labs(subtitle = p2_tit) +
  theme(axis.text = element_text(size = 10), axis.title = element_text(size = 10))

# remove default lines to leave modified ones with alpha value
p2$layers <- p2$layers[-1]

str_overall <- sprintf('%s with (95%%CI: %.2f-%.2f)', fig_roc$ALL$label, pROC::ci(fig_roc$ALL$roc)[1], pROC::ci(fig_roc$ALL$roc)[3])

```

```{r fig3-prep}

tmp <- figs %>% 
  mutate(
    status   = status   %>% factor() %>% forcats::fct_recode('MA-'   = 'NAIVE',      'MA+'    = 'NON-NAIVE'),
    response = response %>% factor() %>% forcats::fct_recode('DOL30' = 'log(V3/V1)', 'DOL128' =  'log(V4/V1)')
  ) %>% 
  select(paired:response, 'train' = training, 'k-fold cv' = cv, 'test: GAM' = test_gam, 'test: PNG' = test_png) %>% 
  gather(col, plot, train, `k-fold cv`, `test: GAM`, `test: PNG`) %>% 
  mutate(col = factor(col, levels = c('train', 'k-fold cv', 'test: GAM', 'test: PNG'))) %>% 
  arrange(status, response)

# selectively add plot title
tmp$plot <- tmp %>% 
  pmap(\(status, response, col, plot, ...) {
    if(col == 'train')
      plot + labs(title = sprintf('%s%s', status, response))
    else
      plot
  })

```

```{r fig3-main-top}
#| echo: false
#| message: false
#| warning: false
#| fig-height: 6.5
#| fig-width:  12
#| column: page-right

# main figure for DOL128
fig_dol128 <- tmp %>% filter(response == 'DOL128') %>% pull(plot)

patchwork::wrap_plots(
  fig_dol128[[1]], fig_dol128[[2]], fig_dol128[[3]], fig_dol128[[4]], p1,
  fig_dol128[[5]], fig_dol128[[6]], fig_dol128[[7]], fig_dol128[[8]], p2,
  nrow = 2
) + patchwork::plot_annotation(tag_levels = 'a')

```

```{r get-models}
source('code/utils/helpers_diablo.R')

# load models
our_models <- readr::read_rds('results/other_outputs/trained_models.rds')

# fix comps
our_models <- our_models %>% mutate(comp = gsub('comp', '', comp) %>% as.numeric())

# fix labels
our_models <- our_models %>% 
  mutate(
    status   = status   %>% factor() %>% forcats::fct_recode('MA-' = 'NAIVE', 'MA+' = 'NON-NAIVE'),
    response = response %>% factor() %>% forcats::fct_recode('DOL30' = 'log(V3/V1)', 'DOL128' = 'log(V4/V1)'),
    model = sprintf('%s %s', response, status)
)

our_models <- our_models %>% 
  pmap(function(status, response, model, diablo, comp, ...) {
    x <- extract_loadings(diablo) %>% filter(comp <= !!comp)
    tibble(status, response, model, features = list(x))
  }) %>% bind_rows()

```

```{r upset}
#| warning: false
#| message: false

library(tidyverse)
library(ComplexUpset)

# order of omics
blocks <- c('flow_type', 'luminex_cytokines', 'proteomics', 'metabolomics', 'epigenetics', 'transcriptomics')

my_upset <- our_models %>%
  mutate(model = sprintf('%s%s', status, response)) %>% 
  unnest(features) %>%
  mutate(block = factor(block, levels = rev(blocks), labels = rev(blocks))) %>%
  rename(feature = block) %>%
  select(-status, -response, -comp, -coefficient) %>% distinct() %>% 
  pivot_wider(names_from = model, names_expand = T, values_from = hgnc_symbol, values_fill = NA) %>%
  mutate(across(starts_with('MA'), ~ !is.na(.))) %>% 
  
  upset(
    intersect = c('MA-DOL30', 'MA-DOL128', 'MA+DOL30', 'MA+DOL128'),
    guides = 'over',
    stripes='white',
    min_size = 3,
    width_ratio = 0.4,
    height_ratio = 0.4,
    themes = list(),

    matrix = (
      intersection_matrix(geom = geom_point(size = 2)) +
        scale_y_discrete(position = 'left') +
        theme(
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 7),
        )
    ),

    set_sizes = (
      upset_set_size(position = 'left') +
        geom_text(aes(label = after_stat(count)), size = 3, hjust = -0.1, stat = 'count', colour = 'white') +
        ylab('set size') +
        theme(
          axis.title.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.y = element_blank(),
          axis.line.y = element_blank(),
          axis.text.x = element_text(size = 7)
          )
    ),
    
    annotations = list(
      'composition'=(
        ggplot(mapping = aes(fill = feature)) + 
          geom_bar(stat = 'count', position = 'fill') + 
          scale_y_continuous(labels = scales::percent_format(), position = 'left') +
          ggsci::scale_fill_npg(name = '') +
          theme(
            axis.title.y = element_blank(),
            axis.line = element_blank(),
            axis.ticks = element_blank(),
            axis.text.x = element_blank(),
            axis.text.y = element_text(size = 8), 
            legend.text = element_text(size = 8)
          )
      )
    ),
    
    base_annotations = list(
      'intersection size' = intersection_size(
        text = list(size = 3, check_overlap = F),
        counts = T
      ) +
        scale_y_continuous(position = 'left') +
        theme(
          axis.title.y = element_blank(),
          axis.ticks.x = element_blank(),
          axis.line = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 8)
        )
    ),
    
    queries = list(upset_query(fill = 'red', color = 'red', intersect = c('MA-DOL128', 'MA-DOL30')))
    
  ) & theme(legend.margin = margin(t = 3.5, r = 1.5, b = 3.5, l = 1.5, 'point'))

# legend text on the left 
my_upset <- my_upset & guides(fill = guide_legend(title.hjust = 1, label.hjust = 1, label.position = 'left', size = 0.1))

cmat <- our_models %>%
  mutate(features = set_names(features, model)) %>%
  pull(features) %>%
  map('feature_name') %>%
  ComplexHeatmap::make_comb_mat()

```

```{r sigora}

f <- list(
  'results/tables/sigora_result_min_logV3V1_MApos.tsv',
  'results/tables/sigora_result_min_logV4V1_MApos.tsv',
  'results/tables/sigora_result_min_logV3V1_MAneg.tsv',
  'results/tables/sigora_result_min_logV4V1_MAneg.tsv'
)

df <- f %>% 
  set_names(gsub('^.+_(.+_.+).tsv$', '\\1', f)) %>% 
  map(readr::read_tsv) %>% 
  bind_rows(.id = 'model')

df$model <- df$model %>% 
  factor(
    levels = c('logV3V1_MAneg', 'logV4V1_MAneg', 'logV3V1_MApos', 'logV4V1_MApos'), 
    labels = c('DOL30 MA-',     'DOL128 MA-',    'DOL30 MA+',     'DOL128 MA+')
  )

# create weighing including model coefficients

# fix duplicates from epi
our_models$features <- our_models$features %>% 
  map(~ {
    .x %>% 
      mutate(hgnc_symbol = stringr::str_split(hgnc_symbol, ';')) %>%
      unnest(hgnc_symbol) %>% 
      distinct()
  })

our_models <- our_models %>% 
  select(model, features) %>% 
  unnest(features)

# arbitrary sum of coefficients...
ff <- df %>%
  select(model, pathway_id, description, gene_ratio, pvalue, bonferroni, cd_genes) %>%
  mutate(cd_genes = stringr::str_split(cd_genes, ';')) 

ff$dist <- ff %>%
  pmap(function(model, cd_genes, ...) {
    cd_genes %>%
      tibble(hgnc_symbol = .) %>% 
      filter(hgnc_symbol != '') %>% 
      mutate(model = !!model) %>%
      left_join(our_models, by = c('hgnc_symbol', 'model')) %>%
      select(feature_name, hgnc_symbol, coefficient) %>%
      na.omit()
  })

ff$score <- ff$dist %>% map(pull, coefficient) %>% map(~ . ^ 2) %>% map_dbl(sum) %>% scales::rescale()

gg <- ff %>% 
  select(model, rows = description, gene_ratio) %>% 
  spread(model, gene_ratio, fill = 0) %>% 
  column_to_rownames('rows') %>% 
  as.matrix()

# heatmap(gg)
keep <- apply(gg, 1, function(x) sum(x != 0)) > 1
gg <- gg[keep, ]

rows <- gg         %>% dist() %>% hclust()
cols <- gg %>% t() %>% dist() %>% hclust()

keep <- df %>% 
  group_by(description) %>% 
  slice_min(pvalue, n = 1) %>% 
  ungroup() %>% 
  slice_min(bonferroni, n = 50) %>% 
  pull(description) %>% 
  unique()

hh <- gg %>%
  as_tibble(rownames = 'description') %>% 
  gather(model, gene_ratio, starts_with('DOL')) %>%
  
  # select pathways only...
  filter(description %in% keep) %>%
  
  left_join(select(ff, model, pathway_id, description, bonferroni, score)) %>%
  
  # reorder based on clustering
  mutate(
    gene_ratio = replace_na(gene_ratio, 0), bonferroni = replace_na(bonferroni, 1), score = replace_na(score, 0),
    description = factor(description, levels = rows$labels[rev(rows$order)]), 
    model = factor(model, levels = cols$labels[rev(cols$order)])
  ) %>% 
  
  # split model into response/status to plot
  mutate(split = model) %>% 
  separate(split, into = c('response', 'status'), sep = ' ') %>% 
  mutate(response = factor(response, levels = c('DOL30', 'DOL128')))

# plot

# bold modules
m <- c(
  "plain", "plain", "plain", "plain", 
  "plain", "plain", "plain", "plain", 
  "plain", "plain", "plain", "plain", 
  "plain", "plain", "plain", "plain", 
  "plain", "plain", "bold",  "plain",
  "plain", "bold",  "bold",  "plain",
  "plain", "plain", "plain", "plain", 
  "plain", "plain", "bold",  "plain"
)

sigora <- hh %>% 
  
  # remove points when no overlap exists between pathway and model features + NetworkAnalyst interactors
  mutate(score = ifelse(gene_ratio == 0, NA, score)) %>% 
  
  ggplot(aes(x = response, y = description, size = score, fill = gene_ratio)) +
  geom_point(aes(alpha = bonferroni > 0.05), pch = 21, colour = 'white') +
  
  scale_x_discrete(position = 'top') +
  scale_y_discrete(position = 'right') +
  
  scale_size_continuous(range = c(1.5, 6.5), breaks = c(0, 0.05, 0.5)) +
  scale_alpha_discrete(range = c(1, 0.25), guide = 'none') +
  scale_fill_viridis_c(option = 'F', breaks = seq(0, max(hh$gene_ratio), by = 0.05), labels = c('0.0', '', '', '', '0.2', '', '', '', '0.4', '', '', '', '0.6'), na.value = 'lightgrey') +
  
  facet_wrap(~ status, scales = 'free_x') +
  
  guides(
    size   = guide_legend(title = 'aggregate\nscore', position = 'right', override.aes = list(colour = 'black')),
    fill   = guide_colorsteps(title = 'gene\nratio', ),
    # colour = guide_colorsteps(title = 'gene\nratio'),
    shape  = guide_none()
  ) +
  
  theme(
    strip.placement = 'outside',
    strip.text = element_text(hjust = 0.5, size = 10, face = 'bold'),
    axis.text = element_text(size = 10),
    axis.text.y = element_text(face = m),
    axis.title = element_blank()
  )

```

```{r gsea-direction}

# include direction estimates from gsea at dol0
gsea <- readr::read_rds('results/other_outputs/gsea_tables.rds')

direction <- gsea %>% 
  select(site, query, formula, gsea) %>% 
  filter(site == 'GAM', grepl('^MA', query)) %>%
  unnest(gsea) %>% 
  rename(description = pathway) %>% 
  filter(description %in% hh$description) %>% 
  select(query, description, NES) %>% 
  separate(query, into = c('status', 'response'), sep = 3)

# maintain order of the pathways from above after merging...
l <- levels(hh$description)

pp <- hh %>% 
  select(description, status, response, gene_ratio, bonferroni, score) %>%
  
  # add NES direction
  left_join(direction) %>% mutate(dir = factor(sign(NES), levels = c(-1, 1), labels = c('neg', 'pos'))) %>%

  # remove points when no overlap exists between pathway and model features + NetworkAnalyst interactors
  mutate(score = ifelse(gene_ratio == 0, NA, score)) %>% 
  
  # re-order as before
  mutate(description = factor(description, levels = l), response = factor(response, levels = c('DOL30', 'DOL128')))

# pp %>% 
sigora <- pp %>%
  ggplot(aes(x = response, y = description, size = score, fill = gene_ratio)) +
  geom_point(aes(shape = factor(dir), alpha = bonferroni > 0.05), colour = 'white') +
  
  scale_shape_manual(values = c(25, 24)) +
  scale_x_discrete(position = 'top') +
  scale_y_discrete(position = 'right') +
  scale_alpha_discrete(range = c(1, 0.25), guide = 'none') +
  scale_size_continuous(range = c(2, 8), breaks = c(0, 0.05, 0.5)) +
  scale_fill_viridis_c(option = 'F', breaks = seq(0, max(hh$gene_ratio), by = 0.05), labels = c('0.0', '', '', '', '0.2', '', '', '', '0.4', '', '', '', '0.6'), na.value = 'lightgrey') +

  facet_wrap(~ status, scales = 'free_x') +
  
  guides(
    shape  = guide_legend(title = 'direction', position = 'right', reverse = T, override.aes = list(colour = 'black', size = 4), order = 1),
    size   = guide_legend(title = 'score', position = 'right', override.aes = list(colour = 'black'), order = 2),
    fill   = guide_colorsteps(title = 'gene\nratio', order = 3)
    # shape  = guide_none()
  ) +
  
  theme(
    strip.placement = 'outside',
    strip.text = element_text(hjust = 0.5, size = 10, face = 'bold'),
    axis.title = element_blank(),
    axis.text = element_text(size = 10),
    axis.text.y = element_text(face = m)
  )

```

```{r fig3-main-bottom}
#| echo: false
#| message: false
#| warning: false
#| fig-height: 6.5
#| fig-width:  12
#| column: page-right

patchwork::wrap_plots(
  my_upset + theme(axis.text = element_text(size = 9)),
  patchwork::free(sigora + scale_y_discrete(position = 'left') + theme(axis.text = element_text(size = 9))),
  widths = c(1, 1.75),
  nrow = 1
) + 
  patchwork::plot_annotation(tag_levels = list(c('k', 'l', '', '', 'm'))) &
  theme(
    # legend.key.size = unit(0.01, 'npc'),
    legend.key.width = unit(0.01, 'npc'),
    legend.key.height = unit(0.03, 'npc'),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10)
  )

```

**Figure 3: Discovery and cross-cohort validation of multi-omic signatures predicting HBV immunogenicity in MA– and MA+ newborns**

::: {.callout-tip collapse="true" title="Overall test classification performance"}
When classifying neonates with HBsAb response above the mean after the full series was: ALL (n=102) AUC=0.67; p=0.002 with (95%CI: 0.56-0.77).
:::

```{r fig3-supp-2}
#| echo: false
#| message: false
#| warning: false
#| fig-height: 6.5
#| fig-width:  9.6
#| column: page-right

# supp figure for DOL30
fig_dol30 <- tmp %>% filter(response == 'DOL30') %>% pull(plot)

patchwork::wrap_plots(
  fig_dol30[[1]], fig_dol30[[2]], fig_dol30[[3]], fig_dol30[[4]],
  fig_dol30[[5]], fig_dol30[[6]], fig_dol30[[7]], fig_dol30[[8]],
  nrow = 2
) + patchwork::plot_annotation(tag_levels = 'a')

```

**Supplemental Figure 5: Discovery and cross-cohort validation of multi-omic predictors for early (DOL30) HBV immunogenicity**

```{r fig3-supp-3}
#| echo: false
#| message: false
#| warning: false
#| fig-height: 3.5
#| fig-width:  9.6
#| column: page-right

# hack
applyto <- c('MA-', 'MA+')
names(applyto) <- rev(applyto)

figs %>% 
  mutate(
    status   = status   %>% factor() %>% forcats::fct_recode('MA-'   = 'NAIVE',      'MA+'    = 'NON-NAIVE'),
    response = response %>% factor() %>% forcats::fct_recode('DOL30' = 'log(V3/V1)', 'DOL128' =  'log(V4/V1)')
  ) %>% 
  select(paired:response, plot = discord) %>% 
  arrange(status, response) %>%
  pmap(function(status, response, plot, ...) {
    plot + 
      labs(
        title =    sprintf('%s %s', status, response),
        subtitle = sprintf('applied to %s neonates', applyto[as.character(status)])
      )
  }) %>% patchwork::wrap_plots(nrow = 1) + patchwork::plot_annotation(tag_levels = 'a')

```

**Supplemental Figure 6: Distinct baseline molecular correlates of HBV immunogenicity in MA– and MA+ newborns**

```{r fig3-supp-1-setup}
#| include: false

library(tidyverse)
ggplot2::theme_set(
  cowplot::theme_cowplot() + 
    ggplot2::theme(
      strip.background = ggplot2::element_blank(),
      axis.title    = element_text(size = 10),
      plot.title    = element_text(face = 'bold', size = 10),
      plot.subtitle = element_text(size = 10)
    )
)

# helper
mystars <- function(p.value) {
  unclass(
    symnum(
      p.value, 
      corr = FALSE, 
      na = FALSE, 
      cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1), 
      symbols = c("***", "**", "*", "†", "")
    )
  )
}

# helper
my_formats <- list(cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1), symbols = c("***", "**", "*", "†", "ns"))

palette <- ggsci::pal_npg()(2)

# name the neonates by HBsAb status at birth...
status_names <- c('HBsAb MA-', 'HBsAb MA+')

# order of omics
block_names <- c('multiomics', 'epigenetics', 'transcriptomics', 'proteomics', 'luminex_cytokines', 'metabolomics', 'flow_type')

```

```{r fig3-supp-1-load}
#| include: false

a <- readr::read_rds(file = 'results/other_outputs/glmnet_variablecv_top20p_results_summarized_rho.rds')
b <- readr::read_rds(file = 'results/other_outputs/diablo_variablecv_top20p_results_summarized_rho.rds')

# nest params and subsets
a <- a %>% nest(params = c(paired, alpha:n))
b <- b %>% nest(params = c(paired, comp, ncomp:n)) %>% mutate(block = 'multiomics')

# combine + hyperparameter search
topmodels <- bind_rows(a, b) %>%
  filter(cv_type == 'kfold', grp == 'ALL') %>%
  group_by(across(dol:response), block, type) %>% 
  arrange(desc(cvm)) %>% 
  slice(1) %>%
  ungroup()

topmodels_dol0 <- topmodels %>% filter(dol == 0)

```

```{r fig3-supp-1-single-multi}
#| include: false

# fix factors
tmp <- topmodels_dol0 %>% 
  unite(id, dol:response) %>%
  mutate(block = factor(block, levels = block_names)) %>% 
  mutate(block = forcats::fct_rev(block)) %>% 
  arrange(block, cvm)

tmp_stat <- ggpubr::compare_means(cvm ~ block, data = tmp, ref.group = 'multiomics', paired = T) %>% rename(block = group2)

fig3a <- tmp %>% 
  mutate(
    block = factor(block, levels = block_names),
    block = forcats::fct_rev(block)
  ) %>% 
  ggplot(aes(cvm, block)) + 
  ggridges::geom_density_ridges(aes(colour = block, fill = block), alpha = 0.25, bandwidth = 0.05, quantile_lines = T, quantiles = 2) +
  geom_vline(xintercept = 0) +
  geom_text(data = tmp_stat, aes(x = 0.5, label = p.signif), nudge_y = 0.5) +
  annotate("rect", xmin = -Inf, xmax = 0, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = 'dimgrey') + #geom_vline(xintercept = 0) +
  ggsci::scale_color_npg() +
  ggsci::scale_fill_npg() +
  scale_x_continuous(limits = c(-0.2, 0.6)) +
  scale_y_discrete(expand = ggplot2::expansion(add = c(0, 2))) +
  theme(
    legend.position = 'none',
    axis.text.y  = element_text(face = c(rep('plain', 6), 'bold')),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  ) +
  labs(x = "Spearman's rho")

fig3a

```

```{r fig3-supp-1-pre-post}
#| include: false

tmp <- topmodels %>%
  filter(daygrp != 'ALL') %>% 
  unite(id, daygrp:response) %>%
  select(id, dol, block, cvm) %>% 
  mutate(dol = factor(dol == 0, levels = c(F, T), labels = c('post', 'pre')))

tmp_stat <- tmp %>% 
  nest(data = -block) %>% 
  mutate(test = map(data, ~ ggpubr::compare_means(cvm ~ dol, data = ., ref.group = 'pre', paired = T))) %>% 
  unnest(test)

tmp <- tmp %>% 
  spread(dol, cvm) %>% 
  mutate(delta = post - pre) %>% 
  mutate(block = factor(block, levels = block_names)) %>% 
  mutate(block = forcats::fct_rev(block))

fig3b <- tmp %>%
  ggplot(aes(delta, block)) + 
  ggridges::geom_density_ridges(aes(colour = block, fill = block), alpha = 0.1, quantile_lines = T, quantiles = 2) +
  geom_vline(xintercept = 0) +
  geom_text(data = tmp_stat, aes(x = 0.5, label = p.signif), nudge_y = 0.5) +
  annotate("rect", xmin = -Inf, xmax = 0, ymin = -Inf, ymax = Inf, alpha = .1, fill = 'dimgrey') +
  ggsci::scale_color_npg() +
  ggsci::scale_fill_npg() +
  scale_x_continuous(limits = c(-0.6, 0.6)) +
  scale_y_discrete(expand = ggplot2::expansion(add = c(0, 2))) +
  theme(
    legend.position = 'none',
    axis.text.y  = element_text(face = c(rep('plain', 6), 'bold')),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  ) +
  # labs(x = expression(paste(Delta, rho, " (post - pre)", sep = '')))
  labs(x = 'model improvement\n(post-pre)')

fig3b

custom_labeller <- function(value) {
  gsub("_", "\n", value)
}

figS3c <- tmp %>% 
  gather(time, perf, pre, post) %>% 
  mutate(time = factor(time, levels = c('pre', 'post')), dol = gsub('^(DOL.).+', '\\1', id)) %>% 
  ggplot(aes(time, perf)) + 
  facet_wrap(~block, labeller = labeller(block = custom_labeller), nrow = 1) +
  # facet_grid(dol ~ block) + 
  ggrain::geom_rain(
    # rain.side = 'f2x2', 
    id.long.var = 'id', 
    boxplot.args= list(outlier.colour = NA, show.legend = F),
    violin.args = list(aes(fill = block, colour = block), alpha = 0.25, trim = F, show.legend = F), 
    point.args  = list(pch = 21, show.legend = F),
    line.args   = list(size = 0.1, show.legend = F)
  ) +
  ggsignif::geom_signif(
    test = 'wilcox.test',
    # y_position = 0.35,
    comparisons = list(c('pre', 'post')),
    map_signif_level = function(x) paste("p =", scales::pvalue(x)), test.args = list(paired = T, exact = F)
  ) +
  ggsci::scale_fill_npg() +
  ggsci::scale_color_npg() +
  labs(y = "Spearman's rho") +
  theme(strip.text = element_text(size = 8))

figS3c

```

```{r fig3-supp-1-constraint}
#| include: false

b <- readr::read_rds(file = 'results/other_outputs/diablo_variablecv_top20p_results_summarized_rho.rds')

# nest params and subsets
b <- b %>% nest(params = c(paired, comp, nkeep, n)) %>% mutate(block = 'multiomics')

# combine + hyperparameter search
topmodels_diablo <- b %>%
  filter(cv_type == 'kfold', grp == 'ALL') %>%
  mutate(block = factor(block, levels = block_names)) %>% 
  group_by(across(dol:response), block, type, design) %>% 
  arrange(desc(cvm)) %>% 
  slice(1) %>%
  ungroup()

fig3c <- topmodels_diablo %>% 
  filter(dol == 0) %>% 
  unite(id, dol:response) %>%
  unnest(params) %>% 
  
  # create dummy variable for design
  mutate(design = design %>% map(as.character) %>% map(unique) %>% map_chr(last) %>% map_dbl(as.numeric) %>% '!='(0) %>% factor(labels = c('no', 'yes'))) %>% 
  
  # for each model, select best constrained, unconstrained
  group_by(id, design) %>% 
  arrange(desc(cvm)) %>% 
  slice(1) %>%
  ungroup() %>% 
  
  # plot
  ggplot(aes(design, cvm)) +

  ggrain::geom_rain(
    rain.side = 'f1x1', 
    id.long.var = 'id', 
    boxplot.args= list(show.legend = F),
    violin.args = list(aes(fill = design, colour = design), alpha = 0.25, trim = F, show.legend = F), 
    point.args  = list(pch = 21, show.legend = F),
    line.args   = list(size = 0.1, show.legend = F)
    ) +
  
  stat_summary(fun = 'median', colour = 'black', geom = 'point', pch = '_', size = 8) +
  # annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 0, alpha = 0.1, fill = 'dimgrey') + #geom_hline(yintercept = 0) +
  labs(x = 'multi-omic constraint', y = "Spearman's rho") +
  scale_y_continuous(expand = expansion(mult = 0.1)) +
  ggsci::scale_fill_npg() +
  ggsci::scale_color_npg() +
  ggsignif::geom_signif(
    test = 'wilcox.test',
    y_position = 0.35,
    comparisons = list(c('no', 'yes')),
    map_signif_level = function(x) paste("p =", scales::pvalue(x)), test.args = list(paired = T, exact = F)
  )

fig3c

```

```{r fig3-supp-1-overfit}
#| include: false

# helper - summarize
make_perf <- function(result) {
  r <- result %>% split(.$type)
  r$training <- r$training %>% 
    unnest(assess) %>% 
    group_by(comp) %>% 
    summarise(estimate = cor(actual, prediction, method = 'spearman'))
  r$cv <- r$cv %>% 
    unnest(assess) %>% 
    group_by(comp, id) %>% 
    summarise(estimate = cor(actual, prediction, method = 'spearman'), .groups = 'drop_last') %>% 
    summarise(estimate = mean(estimate), .groups = 'drop')  
  r %>% bind_rows(.id = 'type') %>% mutate(comp = gsub('comp', '', comp) %>% as.integer())
}

# read in
df <- readr::read_rds('results/other_outputs/diablo_variablecv_top20p_results_abridged.rds')

# subset
df <- df %>% filter(dol == 0, daygrp == 'DOL7', grp == 'ALL', status != 'ALL')

# extract and summarize
df$gg <- df$results %>% map(make_perf, .progress = T)

df <- df %>% 
  select(dol:response, ncomp:design, gg) %>% 
  unite(model, dol:nkeep) %>% 
  mutate(design = design %>% map(as.vector) %>% map(unique) %>% map(as.character) %>% map_chr(last)) %>% 
  unnest(gg)

figS3a <- df %>% 
  ggplot(aes(comp, estimate, colour = design, group = paste(model, design, type))) +
  geom_line(aes(linetype = type), size = 0.1, show.legend = F) +
  facet_wrap(~ design, nrow = 1) +
  ggsci::scale_color_npg() +
  scale_x_continuous(name = "sPLS-DA component", breaks = c(1:10), limits = c(1, 10)) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(y = "Spearman's rho") +
  theme(legend.position = 'bottom')

figS3b <- df %>% 
  spread(type, estimate) %>%
  mutate(overfit = training - cv) %>%
  ggplot(aes(x = comp, colour = design, fill = design, group = paste(model, design))) +
  geom_ribbon(aes(ymin = cv, ymax = training), size = 0.1, alpha = 0.05, show.legend = F) +
  facet_wrap(~ design, nrow = 1) +
  ggsci::scale_color_npg() +
  ggsci::scale_fill_npg() +
  scale_x_continuous(name = "sPLS-DA component", breaks = c(1:10), limits = c(1, 10)) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(y = "Spearman's rho") +
  theme(legend.position = 'bottom')

```

```{r fig3-supp-1-stratification}
#| include: false

figS4 <- topmodels_dol0 %>%
  mutate(status_grp = factor(status != 'ALL', levels = c(F, T), labels = c('F', 'T'))) %>%
  mutate(status = forcats::fct_recode(status, 'ALL' = 'ALL', 'MA-' = 'NAIVE', 'MA+' = 'NON-NAIVE')) %>%
  mutate(daygrp = forcats::fct_rev(daygrp)) %>%
  filter(daygrp == 'DOL7') %>% 

  ggplot(aes(status_grp, cvm)) +
  
  ggrain::geom_rain(
    rain.side = 'f1x1', 
    boxplot.args= list(outlier.colour = NA, show.legend = F),
    violin.args = list(aes(fill = status_grp, colour = status_grp), alpha = 0.25, trim = F, show.legend = F), 
    point.args  = list(pch = 21, show.legend = F)
  ) +
  
  ggpubr::stat_compare_means(method = 'wilcox', comparisons = list(c('F', 'T')), label = 'p.signif', symnum.args = my_formats) +

  # annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 0, alpha = 0.1, fill = 'dimgrey') + #geom_hline(yintercept = 0) +

  ggsci::scale_color_npg(name = 'stratified') +
  ggsci::scale_fill_npg(name = 'stratified') +

  theme(legend.position = 'bottom') +
  labs(x = 'stratified by matAbs', y = "Spearman's rho")

figS4

```

```{r fig3-supp-1}
#| fig-height: 8
#| fig-width:  11
#| column: page-right

patchwork::wrap_plots(
  patchwork::wrap_plots(fig3a + scale_y_discrete(position = 'right'), figS4, fig3c, nrow = 1, widths = c(1, 1.25, 1.25)),
  figS3a,
  figS3b,
  nrow = 3
) + patchwork::plot_annotation(tag_levels = 'a')

```

**Supplemental Figure 4: Multi-omic integration improves predictive performance and reveals population-specific correlates of HBV immunogenicity**
