---
title: "Figure 2 & Associated Supplemental Figures"
format: 
  html:
    fig-format: svg
    filters:
      - inline-svg
editor: source
author: "C.P. Shannon"
execute:
  echo:    false
  warning: false
  message: false
self-contained: true
self-contained-math: true
---

```{r setup}
#| include: false

library(tidyverse)
ggplot2::theme_set(
  cowplot::theme_cowplot() + 
    ggplot2::theme(
      strip.background = ggplot2::element_blank(),
      axis.title    = element_text(size = 10),
      plot.title    = element_text(face = 'bold', size = 10),
      plot.subtitle = element_text(size = 10)
    )
)

# helper
mystars <- function(p.value) {
  unclass(
    symnum(
      p.value, 
      corr = FALSE, 
      na = FALSE, 
      cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1), 
      symbols = c("***", "**", "*", "†", "")
    )
  )
}

palette <- ggsci::pal_npg()(2)
proj3cols <-  c('#3C5488FF', '#6F286AFF', '#E64B35FF', 'darkgrey')
# proj3cols <-  c('blue', 'purple', 'red', 'grey')

# name the neonates by HBsAb status at birth...
status_names <- c('HBsAb MA-', 'HBsAb MA+')

```

```{r import-data}

# read-in metadata
meta <- readRDS('data/processed/gam/import_metadata.rds') %>% 
  select(subject_id = 'Unique Identifier', grp = 'VaccineGrp', day = DayGrp, sex = Sex) %>% 
  distinct()

# read-in maternal metadata
maternal <- aws.s3::s3read_using(readr::read_csv, object = 's3://epichipc-main/Clean_Data/Clinical/GAM_Main_Tier1_Clean.csv') %>%
# maternal <- readr::read_csv('data/raw/gam/GAM_Main_Tier1_Clean.csv') %>%
  select(
    subject_id    = 'Unique Identifier', 
    birth_weight  = 'Weight (grams) V1', 
    maternal_age  = 'Age (years) Maternal', 
    maternal_eth  = 'Ethnicity Maternal', 
    maternal_race = 'Race Maternal'
  )

# combine
meta <- full_join(meta, maternal)

# read-in and make it long
titers <- aws.s3::s3read_using(readr::read_csv, object = 's3://epichipc-main/Clean_Data/SingleOmics/current/GAMMAIN_SINGLEOMICS_TITERS.csv') %>%
# titers <- readr::read_csv('data/raw/gam/GAMMAIN_SINGLEOMICS_TITERS.csv') %>% 
  select(`Subject ID`, `Visit Num`, ConcentrationImputed) %>% # we're only using imputed titer values
  spread(`Visit Num`, ConcentrationImputed) %>%
  mutate(across(c(MAT, V1, V3, V4), as.numeric)) %>% 
  rename(subject_id = 'Subject ID') %>% 
  gather(visit, titer, MAT, V1, V3, V4)

df <- left_join(meta, titers) %>% mutate(grp = factor(grp, levels = c('Delayed', 'HBV', 'BCG', 'HBV+BCG')))

rm(meta, titers)

```

```{r transform-data}

# log-titers
df <- df %>% mutate(titer = log(titer))

# spread
df <- df %>% spread(visit, titer) 

# clean up factors and add ab status
df <- df %>% 
  # define ab status based on V1 (NOTE: log transformed above!!!)
  mutate(day = factor(day, ordered = T), status = factor(V1 < log(2.5), levels = c(F, T), labels = c('non-naive', 'naive'))) %>%
  select(-MAT)

# pivot long
df <- df %>% pivot_longer(matches('^V|^delta'), names_to = 'visit', values_to = 'titer')

dat <- df %>% 
  mutate(
    grp = factor(grp, levels = c('HBV', 'BCG', 'HBV+BCG')),
    status = factor(status, levels = c('naive', 'non-naive'), labels = status_names),
    day = gsub('DOL', '', day) %>% as.numeric()
  ) %>%
  spread(visit, titer) %>% 
  filter(grp != 'Delayed') %>% 
  na.omit()

```

```{r fig-container}
fig <- list()
```

```{r fig-2a}

tmp <- dat %>% 
  gather(time, titer, V1, V3, V4) %>% 
  mutate(time = factor(time, levels = c('V1', 'V3', 'V4'), labels = c('0', '30', '128')))

tmp <- c('all neonates' = list(tmp), split(tmp, tmp$status))

tmp <- tmp %>% 
  imap(~ {
    .x %>% 
      ggplot(aes(time, exp(titer), colour = time)) +
      geom_path( position = position_jitter(width = 0.25, seed = 123), aes(group = subject_id), colour = 'black', linewidth = 0.1, alpha = 0.25, show.legend = F) +
      geom_point(pch = 21, position = position_jitter(height = 0.05, width = 0.25, seed = 123), aes(group = subject_id), show.legend = F) +
      geom_violin(trim = T, scale = 'width', na.rm = T, width = 0.5, alpha = 0.75, show.legend = F) +
      stat_summary(fun = 'mean', pch = '_', stroke = 1, size = 3, colour = 'black') +
      ggplot2::geom_hline(yintercept = 10, linetype = 'dotted') +
      ggplot2::scale_y_continuous(trans = c("log10"), labels = scales::label_log(), expand = expansion(mult = c(0.05, 0.20))) +
      ggplot2::scale_colour_manual(values = ggsci::pal_npg()(3) %>% rev()) +
      ggpubr::stat_compare_means(
        size = 5,
        paired = T,
        # label.y = 4.5,
        label.y = c(4, 4.5),
        label = 'p.signif',
        comparisons = list(c('0', '30'), c('30', '128'))
      ) +
      expand_limits(y = c(1e0, 1e5)) +
      labs(x = 'DOL', y = 'HBsAb titer (mIU/mL)', title = .y) +
      theme(legend.position = 'bottom')
  })

names(tmp) <- c('2a-all', '2a-naive', '2a-non-naive')

fig <- c(fig, tmp)

```

```{r fig-2b}

tmp <- dat %>% 
  select(-V1) %>% 
  pivot_longer(names_to = 'time', values_to = 'titer', cols = starts_with('V')) %>% 
  mutate(time = factor(time, levels = c('V3', 'V4'), labels = c('30', '128')))

levels(tmp$status) <- levels(tmp$status) %>% gsub('.*(MA.).*', '\\1', .)

tmp <- tmp %>% 
  ggplot(aes(time, exp(titer), colour = status, group = paste(status, time))) +
  geom_point(pch = 21, position = position_jitterdodge(jitter.height = 0.05, jitter.width = 0.5, dodge.width = 0.75)) +
  geom_violin(trim = T, scale = 'width', na.rm = T, width = 0.5, alpha = 0.75, position = position_dodge(width = 0.75)) +
  stat_summary(fun = 'mean', pch = '_', stroke = 1, size = 3, colour = 'black', position = position_dodge(width = 0.75)) +
  ggplot2::geom_hline(yintercept = 10, linetype = 'dotted') +
  ggplot2::scale_y_continuous(trans = c("log10"), labels = scales::label_log(), expand = expansion(mult = c(0.05, 0.20))) + 
  scale_color_manual(values = rev(ggsci::pal_npg()(5))) +
  ggpubr::stat_compare_means(
    size = 5,
    label.y = 4.5,
    label = 'p.signif',
    show.legend = F
  ) +
  theme(legend.position = c(0.05, 0.95), legend.direction = 'horizontal') +
  expand_limits(y = c(1e0, 1e5)) +
  labs(x = 'DOL', y = 'HBsAb titer (mIU/mL)')

# tmp

fig$'2b' <- tmp

```

```{r lme-setup}

foo <- dat %>% 
  gather(time, titer, V3, V4) %>%
  mutate(grp_combo  = factor(grp,  levels = c('HBV', 'BCG', 'HBV+BCG'), labels = c('HBV', 'BCG+', 'BCG+'))) %>%
  mutate(time = factor(time, levels = c('V3', 'V4'), labels = c('DOL30', 'DOL128'))) %>% 
  mutate(sex  = factor(sex,  levels = c('Female', 'Male'), labels = c('F', 'M')))

```

```{r lme-run-without-ab-non-naive}
# lme imports
library(nlme)
library(effects)
library(reghelper)

# filter babies with non-detectable titers at birth
data <- foo %>% 
  mutate(birth_weight = birth_weight/1000) %>% 
  filter(status == 'HBsAb MA-') %>%
  as.data.frame()

models <- list(
  fit1 = nlme::lme(
    data = data,
    random = ~ 1 | subject_id, 
    fixed = titer ~ maternal_age + birth_weight + sex + time + day + grp
  ),
  
  fit2 = nlme::lme(
    data = data,
    random = ~ 1 | subject_id, 
    fixed = titer ~ maternal_age + birth_weight + sex + time + day * grp
  ),
  
  fit3 = nlme::lme(
    data = data,
    random = ~ 1 | subject_id,
    fixed = titer ~ maternal_age + birth_weight + sex + time * day * grp
  )
)
```

```{r lme-plots}
# make some plots

# helper
plot_effect <- function(fit) {
  fit$data %>%
    distinct(sex, grp, time) %>%
    crossing(
      birth_weight = median(fit$data$birth_weight),
      maternal_age = median(fit$data$maternal_age),
      day = seq(1, 7, by = 0.01)
    ) %>%
    bernr::bolker_ci(model = fit, conf_level = 0.95, newdat  = .) %>% 
    mutate(across(all_of(c('pred', 'se', 'ci_l', 'ci_h')), exp))
}

template <- function(data, preds, title) {
  data %>% 
    ggplot(aes(day, exp(titer) - min(titer), group = paste(sex, grp, time))) +
    geom_hline(yintercept = 10, linetype = 'dotted') +
    scale_color_manual(values = rev(ggsci::pal_npg()(6))) +
    ggplot2::scale_x_continuous(limits = c(0.5, 7.5), breaks = c(1, 3, 7)) +
    ggplot2::scale_y_continuous(trans = c("log10"), labels = scales::label_log(), expand = expansion(mult = c(0.05, 0.10))) +
    labs(
      title = sprintf('%s (n=%d)', title, nrow(data)),
      x = 'DOL (follow-up delay)', 
      y = 'DOL30\nHBsAb titer (mIU/mL)'
      )
}

table_1 <- models$fit1 %>% broom.mixed::tidy() %>% filter(effect == 'fixed') %>% mutate(p.value = sprintf('%.1g', p.value, mystars(p.value)))
table_2 <- models$fit2 %>% reghelper::simple_slopes() %>% filter(day == 'sstest') %>% rename(p.value = 'Pr(>|t|)') %>% mutate(p.value = sprintf('%.1g', p.value, mystars(p.value)))

tmp_data <- data %>% filter(time == 'DOL30')
tmp_pred <- models$fit1 %>% plot_effect() %>% filter(sex == 'M', grp == 'BCG', time == 'DOL30') 
fig$`2c-all` <- template(tmp_data, tmp_pred, title = 'All neonates') + 
  coord_cartesian(ylim = c(1, 200)) +
  geom_point(position = position_jitter(height = 0.05, width = 0.5), pch = 21) + 
  annotate(
    geom = 'text', 
    x = 1, y = 150, hjust = 0, 
    label = sprintf('slope = %.2g\np = %s', table_1$estimate[table_1$term == 'day'], table_1$p.value[table_1$term == 'day'])
  )

tmp_data <- data %>% filter(grp == 'HBV', time == 'DOL30')
tmp_pred <- models$fit3 %>% plot_effect() %>% filter(sex == 'M', grp == 'HBV', time == 'DOL30') 
fig$`2c-hbv` <- template(tmp_data, tmp_pred, title = 'HBV at DOL0') + 
  coord_cartesian(ylim = c(1, 200)) +
  geom_line(data = tmp_pred, aes(x = day, y = pred), colour = proj3cols[1]) +
  geom_ribbon(data = tmp_pred, aes(x = day, y = pred, ymin = ci_l, ymax = ci_h), colour = NA, fill = proj3cols[1], alpha = 0.1) +
  geom_point(colour = proj3cols[1], position = position_jitter(height = 0.05, width = 0.5), pch = 21) + 
  annotate(
    geom = 'text', 
    x = 1, y = 150, hjust = 0, 
    label = sprintf('slope = %.2g\np = %s', table_2$`Test Estimate`[table_2$grp == 'HBV'], table_2$p.value[table_2$grp == 'HBV'])
  )

tmp_data <- data %>% filter(grp == 'HBV+BCG', time == 'DOL30')
tmp_pred <- models$fit3 %>% plot_effect() %>% filter(sex == 'M', grp == 'HBV+BCG', time == 'DOL30') 
fig$`2c-hbv-bcg` <- template(tmp_data, tmp_pred, title = 'HBV+BCG at DOL0') + 
  coord_cartesian(ylim = c(1, 200)) +
  geom_line(data = tmp_pred, aes(x = day, y = pred), colour = proj3cols[2]) +
  geom_ribbon(data = tmp_pred, aes(x = day, y = pred, ymin = ci_l, ymax = ci_h), colour = NA, fill = proj3cols[2], alpha = 0.1) +
  geom_point(colour = proj3cols[2], position = position_jitter(height = 0.05, width = 0.5), pch = 21) + 
  annotate(
    geom = 'text', 
    x = 1, y = 150, hjust = 0, 
    label = sprintf('slope = %.2g\np = %s', table_2$`Test Estimate`[table_2$grp == 'HBV+BCG'], table_2$p.value[table_2$grp == 'HBV+BCG'])
  )

tmp_data <- data %>% filter(grp == 'BCG', time == 'DOL30')
tmp_pred <- models$fit3 %>% plot_effect() %>% filter(sex == 'M', grp == 'BCG', time == 'DOL30') 
fig$`2c-bcg` <- template(tmp_data, tmp_pred, title = 'BCG at DOL0') + 
  coord_cartesian(ylim = c(1, 200)) +
  geom_line(data = tmp_pred, aes(x = day, y = pred), colour = proj3cols[3]) +
  geom_ribbon(data = tmp_pred, aes(x = day, y = pred, ymin = ci_l, ymax = ci_h), colour = NA, fill = proj3cols[3], alpha = 0.1) +
  geom_point(colour = proj3cols[3], position = position_jitter(height = 0.05, width = 0.5), pch = 21) + 
  annotate(
    geom = 'text', 
    x = 1, y = 150, hjust = 0, 
    label = sprintf('slope = %.2g\np = %s', table_2$`Test Estimate`[table_2$grp == 'BCG'], table_2$p.value[table_2$grp == 'BCG'])
  )

my_comparisons <- list(c('BCG', 'HBV'))

y_pos <- log10(150)

tmp <- dat %>% 
  filter(day == 7, status == 'HBsAb MA-') %>% 
  mutate(titer = exp(V3), grp = grp %>% factor(levels = c('HBV', 'HBV+BCG', 'BCG')))

fig$`2d` <- tmp %>% 
  ggplot(aes(grp, titer, colour = grp)) +
  geom_point(position = position_jitter(height = 0.05, width = 0.1), pch = 21, show.legend = F) + 
  stat_summary(fun.data = 'mean_cl_boot', position = position_nudge(x = 0.25), size = 0.5, linewidth = 0.5, show.legend = F) +
  geom_hline(yintercept = 10, linetype = 'dotted') +
  ggsignif::geom_signif(y_position = y_pos, colour = 'black', size = 0.3, comparisons = my_comparisons, map_signif_level = function(x) paste("p =", scales::pvalue(x)), test.args = list(exact = F)) +
  ggplot2::scale_y_continuous(trans = c("log10"), labels = scales::label_log(), limits = c(1, 200), expand = expansion(mult = c(0.05, 0.10))) +
  scale_colour_manual(values = proj3cols)+
  labs(x = 'vaccine(s) given at DOL0', y = 'DOL30\nHBsAb titer (mIU/mL)', title = sprintf('DOL7: All (n=%d)', nrow(tmp)))

tmp <- dat %>% 
  filter(sex == 'Female', day == 7, status == 'HBsAb MA-') %>% 
  mutate(titer = exp(V3), grp = grp %>% factor(levels = c('HBV', 'HBV+BCG', 'BCG')))

fig$`2d-female` <- tmp %>% 
  ggplot(aes(grp, titer, colour = grp)) +
  geom_point(shape = 17, position = position_jitter(height = 0.05, width = 0.1), pch = 21, show.legend = F) + 
  stat_summary(fun.data = 'mean_cl_boot', position = position_nudge(x = 0.25), size = 0.5, linewidth = 0.5, show.legend = F) +
  geom_hline(yintercept = 10, linetype = 'dotted') +
  ggsignif::geom_signif(y_position = y_pos, colour = 'black', size = 0.3, comparisons = my_comparisons, map_signif_level = function(x) paste("p =", scales::pvalue(x)), test.args = list(exact = F)) +
  ggplot2::scale_y_continuous(trans = c("log10"), labels = scales::label_log(), limits = c(1, 200), expand = expansion(mult = c(0.05, 0.10))) +
  scale_colour_manual(values = proj3cols)+
  labs(x = 'vaccine(s) given at DOL0', y = 'DOL30\nHBsAb titer (mIU/mL)', title = sprintf('Females (n=%d)', nrow(tmp)))

tmp <- dat %>% 
  filter(sex == 'Male', day == 7, status == 'HBsAb MA-') %>% 
  mutate(titer = exp(V3), grp = grp %>% factor(levels = c('HBV', 'HBV+BCG', 'BCG')))

fig$`2d-male` <- tmp %>% 
  ggplot(aes(grp, titer, colour = grp)) +
  geom_point(shape = 15, position = position_jitter(height = 0.05, width = 0.1), pch = 21, show.legend = F) + 
  stat_summary(fun.data = 'mean_cl_boot', position = position_nudge(x = 0.25), size = 0.5, linewidth = 0.5, show.legend = F) +
  geom_hline(yintercept = 10, linetype = 'dotted') +
  ggsignif::geom_signif(y_position = y_pos, colour = 'black', size = 0.3, comparisons = my_comparisons, map_signif_level = function(x) paste("p =", scales::pvalue(x)), test.args = list(exact = F)) +
  ggplot2::scale_y_continuous(trans = c("log10"), labels = scales::label_log(), limits = c(1, 200), expand = expansion(mult = c(0.05, 0.10))) +
  scale_colour_manual(values = proj3cols)+
  labs(x = 'vaccine(s) given at DOL0', y = 'DOL30\nHBsAb titer (mIU/mL)', title = sprintf('Males (n=%d)', nrow(tmp)))

```

<!-- # Figure 2 -->
```{r}
#| label: figure-2
#| fig-height: 9.75
#| fig-width:  8.25
#| column: page-right

interaction_pval <- models$fit2 %>% broom.mixed::tidy() %>% filter(term == 'day:grpBCG') %>% pull(p.value)
interaction <- fig$`2c-hbv` + labs(subtitle = sprintf('delay by vaccine(s) given at DOL0 group interaction, p* = %.2g', interaction_pval)) + theme(hjust = 0, plot.subtitle = element_text(size = 12, face = 'italic'))

patchwork::wrap_plots(
  patchwork::wrap_plots(fig$`2a-naive`, fig$`2a-non-naive`, fig$`2b`, axis_titles = 'collect', nrow = 1),
  patchwork::wrap_plots(interaction, fig$`2c-hbv-bcg`, fig$`2c-bcg`, axis_titles = 'collect', nrow = 1),
  patchwork::wrap_plots(fig$`2d`, fig$`2d-female`, fig$`2d-male`,     axis_titles = 'collect', nrow = 1),
  nrow = 3
) + 
  patchwork::plot_annotation(tag_levels = 'a')

```

**Figure 2: Maternal antibodies, as well as the timing and sequence of immunization with HBV and, BCG and OPV, impact HBV vaccination**

<!-- # Figure S2 -->
```{r}
#| label: figure-S2
#| fig-height: 4.875
#| fig-width:  8.25
#| column: page-right

# filter babies with non-detectable titers at birth
data <- foo %>% 
  mutate(birth_weight = birth_weight/1000) %>% 
  filter(status == 'HBsAb MA-') %>%
  as.data.frame()

data %>% 
  mutate(grp = factor(grp, levels = c('HBV', 'HBV+BCG', 'BCG'))) %>% 
  ggplot(aes(day, exp(titer-min(titer)))) +
  geom_point(aes(colour = time, shape = time), position = position_jitter(width = 0.3)) +
  # ggpubr::stat_regline_equation(aes(group = time), colour = 'black', label.y = c(4.75, 5)) +
  stat_smooth(method = 'lm', aes(group = time), colour = 'black', alpha = 0.1, linewidth = 0.5) +
  scale_shape(solid = F) +
  ggsci::scale_color_npg(name = 'time') +
  ggplot2::scale_x_continuous(limits = c(0.5, 7.5), breaks = c(1, 3, 7)) +
  ggplot2::scale_y_continuous(trans = c("log10"), labels = scales::label_log(), expand = expansion(mult = c(0.05, 0.10))) +
  facet_grid(. ~ grp, scales = 'free') +
  labs(y = 'HBsAb log(mIU/mL)') +
  theme(legend.position = 'bottom')

```

**Supplemental Figure X: Association between HBsAb titers, timing and sequence of immunization with HBV and, BCG and OPV, after a single-dose (DOL30) or the full-series (DOL128)**

<!-- # Figure S3 -->
```{r}

# cor plots titers
tmp <- dat
tmp <- split(tmp, tmp$status)

tmp <- tmp %>% 
  imap(~ {
    .x %>% 
      ggplot(aes(exp(V3), exp(V4))) + 
      geom_point(pch = 21, position = position_jitter(width = 0.025)) +
      stat_smooth(method = MASS::rlm, alpha = 0.1, size = 0.5, colour = 'black', fullrange = T) +
      ggpubr::stat_cor(output.type = 'text', method = 'spearman', label.x.npc = 1, label.y.npc = 0.05, hjust = 1, aes(label = sprintf('rho=%.2f (p=%.2g)', after_stat(r), after_stat(p)))) +
      ggplot2::scale_x_continuous(trans = c("log10"), labels = scales::label_log()) +
      ggplot2::scale_y_continuous(trans = c("log10"), labels = scales::label_log()) +
      labs(x = 'DOL30\nHBsAb titer (mIU/mL)', y = 'DOL128\nHBsAb titer (mIU/mL)', title = .y)
  })

names(tmp) <- c('s2-naive', 's2-non-naive')
fig <- c(fig, tmp)

# cor plots logVX/V1
tmp <- dat %>% mutate(V3 = V3 - V1, V4 = V4 - V1)
tmp <- split(tmp, tmp$status)

tmp <- tmp %>% 
  imap(~ {
    .x %>% 
      ggplot(aes(exp(V3), exp(V4))) + 
      geom_point(pch = 21, position = position_jitter(width = 0.025)) +
      stat_smooth(method = MASS::rlm, alpha = 0.1, size = 0.5, colour = 'black', fullrange = T) +
      ggpubr::stat_cor(output.type = 'text', method = 'spearman', label.x.npc = 1, label.y.npc = 0.05, hjust = 1, aes(label = sprintf('rho=%.2f (p=%.2g)', after_stat(r), after_stat(p)))) +
      ggplot2::scale_x_continuous(trans = c("log10"), labels = scales::label_log()) +
      ggplot2::scale_y_continuous(trans = c("log10"), labels = scales::label_log()) +
      labs(x = 'DOL30\nHBsAb titer (mIU/mL)', y = 'DOL128\nHBsAb titer (mIU/mL)', title = .y, subtitle = 'adj. baseline')
  })

names(tmp) <- c('s2-adj-naive', 's2-adj-non-naive')

fig <- c(fig, tmp)

```

```{r}
#| label: figure-S3
#| fig-height: 3.5
#| fig-width:  11
#| column: page-right

patchwork::wrap_plots(fig$`s2-naive`, fig$`s2-non-naive`, fig$`s2-adj-naive`, fig$`s2-adj-non-naive`, nrow = 1) + patchwork::plot_annotation(tag_levels = 'a')

```

**Supplemental Figure 3: HBsAb titers after a single dose (DOL30) and after a full series (DOL128) are moderately correlated after adjusting for baseline (DOL0) titers**

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-height: 6.5
#| fig-width:  11
#| column: page-right

con <- tribble(
  ~site,    ~data,
  'gambia', readr::read_rds('data/processed/gam/import_responses.rds'),
  'png',    readr::read_rds('data/processed/png/import_responses.rds')
)

df_titers <- con %>%
  unnest(data) %>%
  rename(subject_id = `Subject ID`) %>% 
  mutate(status = factor(V1 < 2.5, levels = c(F, T), labels = c('HBsAb MA+', 'HBsAb MA-'))) %>% 
  select(-MAT) %>% 
  na.omit() %>% 
  gather(time, titer, V1, V3, V4) %>% 
  mutate(time = factor(time, levels = c('V1', 'V3', 'V4'), labels = c('0', '30', '128')))

# re-order for figure - MA- should appear first
df_titers$status <- df_titers$status %>% forcats::fct_rev()

figs_titers <- df_titers %>% 
  split(.$status) %>% 
  map(~ {
    p1 <- .x %>% 
      ggplot(aes(time, titer, colour = time)) +
      geom_path(position = position_jitter(height = 0.05, width = 0.25, seed = 123), aes(colour = time, group = subject_id), colour = 'black', linewidth = 0.1, alpha = 0.5, show.legend = F) +
      geom_point(pch = 21, position = position_jitter(height = 0.05, width = 0.25, seed = 123), aes(group = subject_id), show.legend = F) +
      geom_violin(trim = T, scale = 'width', na.rm = T, width = 0.5, alpha = 0.75, show.legend = F) +
      stat_summary(fun = 'mean', pch = '_', stroke = 1, size = 3, colour = 'black') +
      ggplot2::geom_hline(yintercept = 10, linetype = 'dotted') +
      ggplot2::scale_y_continuous(trans = c("log10"), labels = scales::label_log()) +
      ggplot2::facet_grid(status ~ site) +
      ggplot2::scale_colour_manual(values = ggsci::pal_npg()(3) %>% rev()) +
      ggpubr::stat_compare_means(
        size = 5,
        paired = T,
        label.y = c(4, 4.45),
        label = 'p.signif',
        comparisons = list(c('0', '30'), c('30', '128'))
      ) +
      expand_limits(y = c(1e0, 1e5)) +
      labs(x = 'DOL', y = 'HBsAb titer (mIU/mL)') +
      theme(legend.position = 'bottom')

    p2 <- .x %>% 
      ggplot(aes(site, titer, colour = time)) +
      geom_point(pch = 21, position = position_jitter(height = 0.05, width = 0.25, seed = 123), aes(group = subject_id), show.legend = F) +
      geom_violin(trim = T, scale = 'width', na.rm = T, width = 0.5, alpha = 0.75, show.legend = F) +
      stat_summary(fun = 'mean', pch = '_', stroke = 1, size = 3, colour = 'black') +
      ggplot2::geom_hline(yintercept = 10, linetype = 'dotted') +
      ggplot2::scale_y_continuous(trans = c("log10"), labels = scales::label_log()) +
      ggplot2::facet_grid(status~ time) +
      ggplot2::scale_colour_manual(values = ggsci::pal_npg()(3) %>% rev()) +
      ggpubr::stat_compare_means(
        size = 5,
        label.y = 4.45,
        label = 'p.signif',
        comparisons = list(c('gambia', 'png'))
      ) +
      expand_limits(y = c(1e0, 1e5)) +
      labs(x = 'site', y = 'HBsAb titer (mIU/mL)') +
      theme(legend.position = 'bottom')
    
    list(p1, p2)
  }) %>% flatten()

patchwork::wrap_plots(figs_titers, ncol = 2, nrow = 2) + patchwork::plot_layout(axes = 'collect_x') + patchwork::plot_annotation(tag_levels = 'a')

```
**Supplemental Figure 7: Qualitative differences in HBsAb titer responses to HBV vaccination in The Gambia and PNG**
