---
title: |
        **NEW**  \
        Figure 4 & Associated Supplemental Figures
subtitle: "Functional Validation: assess REACTOME pathways: ~response @DOL30, @DOL128 in MA-/MA+ Neonates, and more..."
format:
  html:
    embed-resources: true
    fig-format: svg
    filters:
      - inline-svg
editor: source
author: "C.P. Shannon"
execute:
  echo:    false
  warning: false
  message: false
self-contained: true
self-contained-math: true
---

```{r setup}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(rstatix)
library(ggpubr)

ggplot2::theme_set(
  cowplot::theme_cowplot() +
    ggplot2::theme(
      strip.background = ggplot2::element_blank(),
      axis.title    = element_text(size = 10),
      plot.title    = element_text(face = 'bold', size = 10),
      plot.subtitle = element_text(size = 10)
    )
)

annot_table <- function(...) {
  rstatix::add_significance(
    ..., 
    cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, 0.1, 1),
    symbols = c("****", "***", "**", "*", ".", "ns")
    )  
}

export_gsea <- function(gsea, msigdb) {
  msigdb %>% 
    select(pathway_id = geneset, pathway = description) %>% 
    right_join(gsea) %>% 
    arrange(pval) %>% 
    mutate(leadingEdge = leadingEdge %>% map_chr(paste, collapse = ';'))
}

my_pal <- c("#E64B35FF", "#4DBBD5FF", "#00A087FF", "#3C5488FF")

targets <- c(
  "Neutrophil degranulation",
  "Interleukin-1 signaling",
  "Platelet activation, signaling and aggregation",
  "Interleukin-4 and Interleukin-13 signaling"
)

```

# Functional Validation

For each of the following pathways of interest:

- REACTOME_NEUTROPHIL_DEGRANULATION
- REACTOME_INTERLEUKIN_1_SIGNALING
- REACTOME_INTERLEUKIN_4_AND_INTERLEUKIN_13_SIGNALING
- REACTOME_PLATELET_ACTIVATION_SIGNALING_AND_AGGREGATION

## Run Analysis
```{r instead}

# source('01_batch_run_gsea.R')
con <- readr::read_rds('results/other_outputs/gsea_tables.rds')
msigdb <- con$msigdb[[1]]

```

# 1. ~Response @DOL0 in MA-

::: callout-important
### Inclusion of PNG?

- no response @dol30
- consistent @dol128

:::

::: callout-tip
### Rationale

Establish the baseline (@dol0) association between our pathways of interest and
vaccine response:

- in MA- neonates used to train our multiomic models
- using Transcriptomic data at DOL0
- @DOL30 (after 1 dose) 
- @DOL128 (after the 3 dose series)

This provides us with a hypothesis including **DIRECTIONALITY**, which we will then
test.

:::

```{r}

con %>% 
  filter(query == 'MA-DOL30') %>% 
  unnest(gsea) %>% 
  filter(pathway %in% targets | grepl('^mod|^MA', pathway)) %>% 
  select(site, pathway, pval, padj, NES, size) %>% 
  annot_table('padj') %>% 
  knitr::kable(digits = 3, caption = '**fgsea** MA-DOL30 @dol0')

con %>% 
  filter(query == 'MA-DOL128') %>% 
  unnest(gsea) %>% 
  filter(pathway %in% targets | grepl('^mod|^MA', pathway)) %>% 
  select(site, pathway, pval, padj, NES, size) %>%
  annot_table('padj') %>% 
  knitr::kable(digits = 3, caption = '**fgsea** MA-DOL128 @dol0')

```

::: callout-note
### Conclusion

The levels of **ALL** of our target pathways:

- REACTOME_NEUTROPHIL_DEGRANULATION
- REACTOME_INTERLEUKIN_1_SIGNALING
- REACTOME_INTERLEUKIN_4_AND_INTERLEUKIN_13_SIGNALING
- REACTOME_PLATELET_ACTIVATION_SIGNALING_AND_AGGREGATION

Are **significantly associated** with response to vaccination, either:

- after **1-dose**
- after the **3-dose series**.

Relationship is always **OPPOSITE** @dol30, @dol128, which is consistent with
*Fig. 5*.

:::

::: callout-warning
### Leading Edge

Obtain:

- Leading edge from GSEA
- NetworkAnalyst output for each model
- Return the overlap to identify the most salient transcriptomic features

:::

```{r}
#| results: asis

f <- list(
  'results/tables/sigora_result_min_logV3V1_MApos.tsv',
  'results/tables/sigora_result_min_logV4V1_MApos.tsv',
  'results/tables/sigora_result_min_logV3V1_MAneg.tsv',
  'results/tables/sigora_result_min_logV4V1_MAneg.tsv'
)

sigora_genes <- f %>%
  set_names(gsub('^.+_(.+_.+).tsv$', '\\1', f)) %>%
  map(readr::read_tsv) %>%
  bind_rows(.id = 'model') %>%
  select(query = model, pathway_id, pathway = description, cd_genes) %>%
  filter(pathway %in% targets) %>% 
  mutate(
    query = factor(query, levels = c('logV3V1_MAneg', 'logV4V1_MAneg', 'logV3V1_MApos', 'logV4V1_MApos'), labels = c('MA-DOL30', 'MA-DOL128', 'MA+DOL30', 'MA+DOL128')),
    cd_genes = cd_genes %>% stringr::str_split(';')
  )

foremost <- con %>% 
  filter(site == 'GAM') %>% 
  filter(query %in% c('MA-DOL30', 'MA-DOL128')) %>% 
  select(query, gsea) %>% 
  unnest(gsea) %>% 
  inner_join(sigora_genes) %>% 
  select(query, pathway, padj, NES, leadingEdge, pathway_id, cd_genes) %>% 
  mutate(foremostEdge = map2(cd_genes, leadingEdge, ~ .y[.y %in% .x]))

# readr::write_rds(foremost, file = 'reviews/foremostEdge.rds')

foremost %>% 
  filter(grepl('DOL30', query), padj < 0.1) %>%
  mutate(foremostEdge = set_names(foremostEdge, pathway)) %>% 
  pull(foremostEdge) %>% 
  iwalk(~{
    cat('**')
    cat(.y)
    cat(':')
    cat('**')
    cat('  \n')
    cat(paste0(.x, collapse = ', '))
    cat('  \n')
    cat('  \n')
  })

```

# 2. ~DOL1 [BCG, HBV, HBV+BCG] and DOL1 [Delayed]

::: callout-important
### PNG

- No DOL1 in PNG

:::

::: callout-tip
### Rationale

This is "necessary, but not suffient..." The reasoning goes:

- If the levels of these pathways at baseline are important to HBV response
- We would expect them to be modified by vaccination (...with HBV)
- We look @dol1 since this is the most likely time to observe differences caused by vaccination

:::

```{r}

# note we're comparing delayed to BCG here based on focusing on the contrast
# with largest absolute coefficients above...
con %>% 
  filter(grepl('^dol1', query)) %>% 
  unnest(gsea) %>%
  mutate(query = gsub('dol1_', '', query)) %>% 
  select(site, query, pathway, pval, padj, NES, size) %>%
  annot_table('padj') %>% 
  knitr::kable(digits = 3, caption = '**fgsea** vaccine groups @dol1')

```

# 3. ~DOL7 [BCG] and DOL0 [HBV, HBV+BCG]

::: callout-tip
### Rationale

This is to assess whether the levels of these pathways differ at the time of HBV
administration:

- @dol0 in newborns randomized to receive [HBV, HBV+BCG] at birth
- @dol7 in newborns randomized to receive [BCG] at birth

Note, we also assessed whether this change in pathways differed between males
and females, but this was not the case.

:::

```{r}

con %>% 
  filter(query == 'anyhbv0_bcg7') %>% 
  unnest(gsea) %>% 
  filter(pathway %in% targets | grepl('^mod|^MA', pathway)) %>% 
  select(site, pathway, pval, padj, NES, size) %>%
  annot_table('padj') %>% 
  knitr::kable(digits = 3, caption = '**fgsea** [bcg]@dol7 - [hbv, hbv+bcg]@dol0')

```

::: callout-note
### Conclusion

When comparing neonates randomized to receive:

- HBV at birth [HBV, HBV+BCG]
- HBV delayed to DOL7 [BCG at birth]

Some of **SOME** of our target pathways:

- REACTOME_NEUTROPHIL_DEGRANULATION
- REACTOME_INTERLEUKIN_1_SIGNALING
- REACTOME_INTERLEUKIN_4_AND_INTERLEUKIN_13_SIGNALING

Are **DOWN-REGULATED** by DOL7.

:::

::: callout-warning
### Reasoning

- These immune pathways are putatively important to HBV response
- These immune pathways are modulated by the time they receive HBV
  - this may reflect normal immune ontogeny or the effects of BCG at DOL0
- Given the direction of association identified above at DOL0, the direction of change here would predict better responses in this subset.
- Indeed, neonates randomized to receive BCG first had better response by DOL30 (*Fig. 2*).

:::

# 4. **GSEA**: MA-, MA+ newborns @DOL0
```{r baseline-status, eval=FALSE}
#| fig-height: 5
#| fig-width: 6

# in vivo
a_tbl <- con %>% 
  filter(site == 'GAM', query == 'baseline_status') %>% 
  unnest(gsea) %>%
  select(site, pathway, pval, padj, NES, size)

a <- a_tbl %>% 
  left_join(select(msigdb, pathway = description, pathway_id = geneset)) %>%
  filter(!grepl('^mod', pathway)) %>% 
  mutate(
    dir = factor(sign(NES), levels = c(-1, 1), labels = c('\nMA-\nup-regulated', '\nMA+\nup-regulated')),
    label = ifelse(
      padj < 0.05 & row_number() < 9, 
      sprintf('%s\n%s', pathway_id, stringr::str_wrap(pathway, 25)),
      ''
    ),
    
    padj = cut(
      padj, 
      breaks = c(0, 0.001, 0.01, 0.05, 1), 
      labels = c('<0.001', '<0.01', '<0.05', 'ns')
    )
    
  ) %>%

  ggplot(aes(NES, -log10(pval))) +
  geom_vline(xintercept = 0, linewidth = 0.1) +
  geom_point(aes(colour = padj), size = 4, pch = 20, alpha = 0.5, position = position_jitter(width = 0.2, height = 0.2, seed = 123)) +
  ggrepel::geom_text_repel(aes(label = label), position = position_jitter(width = 0.2, height = 0.2, seed = 123), force = 10, size = 3, max.overlaps = Inf, min.segment.length = 0, lineheight = 0.75) +
  viridis::scale_colour_viridis('FDR', option = 'E', direction = -1, discrete = T) +
  facet_grid(. ~ dir, scales = 'free') +
  ggh4x::facetted_pos_scales(
    x = list(
      scale_x_continuous(breaks = c(-3, -2, -1, 0)),
      scale_x_continuous(breaks = c(0, 1, 2, 3))
    )
  ) +
  theme(
    legend.position = c(0.05, 0.80),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10),
    legend.key.width = unit(0.01, 'npc'),
    legend.key.height = unit(0.005, 'npc'),
    legend.key.spacing = unit(0.005, 'npc'),
    panel.spacing.x = unit(-0.05, 'npc'), 
    strip.text = element_text(face = 'italic', size = 9)
  )

a

```

```{r baseline-status-tbl, eval=FALSE}

tmp <- con %>% 
  filter(site == 'GAM', query == 'baseline_status') %>% 
  pull(gsea) %>% first() %>% 
  export_gsea(msigdb) %>% 
  filter(pval < 0.05)

filepath <- 'results/tables'
filename <- 'pre_vaccine_abstatus_invivo_fgsea_results.csv'

readr::write_csv(
  tmp, 
  file = file.path(filepath, filename)
)

aws.s3::put_object(
  file.path(filepath, filename),
  object = file.path('s3://epichipc-main-dataintegration/Casey', filename)
)

```

# 5. **GSEA**: MA-, MA+ *in vitro* & *in vivo* @post-vaccination
```{r post-vaccine-status-prep}

comparison <- con %>% filter(site %in% c('invivo', 'invitro')) %>% mutate(gsea = set_names(gsea, site))

# create some supp. tables
my_tbls <- comparison %>% 
  filter(query == 'status') %>% 
  select(site, gsea) %>% 
  deframe()

my_mat <- my_tbls %>% 
  map(filter, padj < 0.1) %>%
  map('pathway') %>%
  ComplexHeatmap::make_comb_mat()

inboth <- ComplexHeatmap::extract_comb(my_mat, '11')

# container for plots
plots_abstatus <- list()

```

```{r post-vaccine-status-invivo}

# DOL1 in vivo
plots_abstatus$invivo <- comparison %>% 
  filter(site == 'invivo') %>% 
  pull(gsea) %>%
  first() %>% 
  left_join(select(msigdb, pathway = description, pathway_id = geneset)) %>%
  filter(!grepl('^mod', pathway)) %>% 
  mutate(
    dir = factor(sign(NES), levels = c(-1, 1), labels = c('\nMA-\nup-regulated', '\nMA+\nup-regulated')),
    label = ifelse(
      padj < 0.05 & row_number() <= 5, 
      sprintf('%s\n%s', pathway_id, stringr::str_wrap(pathway, 25)),
      ''
    ),
    
    padj = cut(
      padj, 
      breaks = c(0, 0.001, 0.01, 0.05, 1), 
      labels = c('<0.001', '<0.01', '<0.05', 'ns')
    )
    
  ) %>%

  ggplot(aes(NES, -log10(pval))) +
  geom_vline(xintercept = 0, linewidth = 0.1) +
  geom_point(aes(colour = padj), size = 4, pch = 20, alpha = 0.5, position = position_jitter(width = 0.2, height = 0.2, seed = 123)) +
  ggrepel::geom_text_repel(aes(label = label), position = position_jitter(width = 0.2, height = 0.2, seed = 123), force = 10, size = 3, max.overlaps = Inf, min.segment.length = 0, lineheight = 0.75) +
  viridis::scale_colour_viridis('FDR', option = 'D', direction = -1, discrete = T) +
  facet_grid(. ~ dir, scales = 'free') +
  ggh4x::facetted_pos_scales(
    x = list(
      scale_x_continuous(breaks = c(-3, -2, -1, 0)),
      scale_x_continuous(breaks = c(0, 1, 2, 3))
    )
  ) +
  
  theme(
    legend.position = c(0.05, 0.80),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10),
    legend.key.width = unit(0.01, 'npc'),
    legend.key.height = unit(0.005, 'npc'),
    legend.key.spacing = unit(0.005, 'npc'),
    panel.spacing.x = unit(-0.05, 'npc'), 
    strip.text = element_text(face = 'italic', size = 9)
  ) +
  
  labs(
    title = 'In vivo',
    subtitle = 'REACTOME pathways after stimulation\nwith HBV, BCG, or HBV+BCG'
  )

```

```{r post-vaccine-status-invivo-tbl, eval=FALSE}

tmp <- comparison %>% 
  filter(site == 'invivo') %>% 
  pull(gsea) %>% first() %>% 
  export_gsea(msigdb) %>% 
  filter(pval < 0.05)

filepath <- 'results/tables'
filename <- 'post_vaccine_abstatus_invivo_fgsea_results.csv'

readr::write_csv(
  tmp, 
  file = file.path(filepath, filename)
)

aws.s3::put_object(
  file.path(filepath, filename),
  object = file.path('s3://epichipc-main-dataintegration/Casey', filename)
)

```

```{r post-vaccine-status-invitro}

# post-18h in vitro
plots_abstatus$invitro <- comparison %>% 
  filter(site == 'invitro') %>% 
  pull(gsea) %>%
  first() %>% 
  left_join(select(msigdb, pathway = description, pathway_id = geneset)) %>%
  filter(!grepl('^mod', pathway)) %>% 
  mutate(
    dir = factor(sign(NES), levels = c(-1, 1), labels = c('\nMA-\nup-regulated', '\nMA+\nup-regulated')),
    label = ifelse(
      padj < 0.05 & row_number() <= 5, 
      sprintf('%s\n%s', pathway_id, stringr::str_wrap(pathway, 25)),
      ''
    ),
    
    padj = cut(
      padj, 
      breaks = c(0, 0.001, 0.01, 0.05, 1), 
      labels = c('<0.001', '<0.01', '<0.05', 'ns')
    )
    
  ) %>%

  ggplot(aes(NES, -log10(pval))) +
  geom_vline(xintercept = 0, linewidth = 0.1) +
  geom_point(aes(colour = padj), size = 4, pch = 20, alpha = 0.5, position = position_jitter(width = 0.2, height = 0.2, seed = 123)) +
  ggrepel::geom_text_repel(aes(label = label), position = position_jitter(width = 0.2, height = 0.2, seed = 123), force = 10, size = 3, max.overlaps = Inf, min.segment.length = 0, lineheight = 0.75) +
  viridis::scale_colour_viridis('FDR', option = 'C', direction = -1, discrete = T) +
  facet_grid(. ~ dir, scales = 'free') +
  ggh4x::facetted_pos_scales(
    x = list(
      scale_x_continuous(breaks = c(-3, -2, -1, 0)),
      scale_x_continuous(breaks = c(0, 1, 2, 3))
    )
  ) +
  
  theme(
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10),
    legend.position = c(0.05, 0.80),
    legend.key.width = unit(0.01, 'npc'),
    legend.key.height = unit(0.005, 'npc'),
    legend.key.spacing = unit(0.005, 'npc'),
    panel.spacing.x = unit(-0.05, 'npc'), 
    strip.text = element_text(face = 'italic', size = 9)
  ) +
  
  labs(
    title = 'In vitro',
    subtitle = 'REACTOME pathways after stimulation\nwith HBV, BCG, or HBV+BCG'
  )

```

```{r post-vaccine-status-invitro-tbl, eval=FALSE}

tmp <- comparison %>% 
  filter(site == 'invitro') %>% 
  pull(gsea) %>% first() %>% 
  export_gsea(msigdb) %>% 
  filter(pval < 0.05)

filepath <- 'results/tables'
filename <- 'post_vaccine_abstatus_invitro_fgsea_results.csv'

readr::write_csv(
  tmp, 
  file = file.path(filepath, filename)
)

aws.s3::put_object(
  file.path(filepath, filename),
  object = file.path('s3://epichipc-main-dataintegration/Casey', filename)
)

```

```{r post-vaccine-status-cor}

# plot cor gsea
annotation <- msigdb %>% select(pathway = description, pathway_id = geneset, candidate, target)
annotation$pathway_id[1:4] <- ''

pp <- comparison %>% 
  unnest(gsea) %>% 
  select(site, pathway, NES) %>% 
  unite(id, site) %>% 
  spread(id, NES) %>% 
  rename(invitro = 2, invivo = 3) %>% 
  left_join(annotation)

# create some variables to sort pathways
pp <- pp %>% 
  mutate(
    signif   = pathway %in% inboth,
    siglab   = factor(signif, levels = c(F, T), labels = c('≥0.10', '<0.10')),
    poi      = target & !grepl('^mod', pathway),
    interest = candidate, 
    lab      = ifelse(pathway %in% inboth, sprintf('%s\n%s', pathway_id, stringr::str_wrap(pathway, 35)), NA)
  )

cor <- filter(pp, candidate)
cor <- cor.test(cor$invitro, cor$invivo, method = 'spearman')

plots_abstatus$common <- pp %>% 
  ggplot(aes(invitro, invivo)) +
  
  # points
  geom_point(aes(colour = poi, shape = siglab, alpha = interest | signif)) +
  ggrepel::geom_text_repel(na.rm = T, aes(label = lab, colour = poi), size = 3, max.overlaps = Inf, min.segment.length = 0, lineheight = 0.75, show.legend = F) +
  
  # annotate
  annotate(
    geom = 'text', 
    size = 9/.pt, 
    fontface = 'italic',
    lineheight = 0.85,
    vjust = 'inside', 
    hjust = 'inside',
    x = c(-2.00, 3.50),
    y = c(-2.00, 4.00),
    label = c('MA-\nup-regulated', 'MA+\nup-regulated'), 
  ) +
  
  geom_vline(xintercept = 0, linewidth = 0.1) +
  geom_hline(yintercept = 0, linewidth = 0.1) +
  
  scale_x_continuous(limits = c(-3.0, 4.2)) +
  scale_y_continuous(limits = c(-2.5, 4.5)) +
  # ggpubr::stat_cor(data = filter(pp, interest), method = 'spearman', show.legend = F) +
  
  scale_alpha_discrete(name = 'implicated\nby model', range = c(0.05, 1)) +
  scale_shape_manual(name = 'FDR', values = c(20, 21)) +
  scale_color_manual(name = 'POI', values = c('black', '#E64B35FF')) +

  guides(
    shape  = guide_legend(order = 1), 
    alpha  = guide_legend(order = 2), 
    colour = guide_legend(order = 3), 
  ) +
  
  labs(
    x = 'NES in vitro',
    y = 'NES in vivo',
    title = 'In vivo vs. in vitro',
    subtitle = sprintf('R = %.2f, p = %.2g\nfor candidate REACTOME pathways', cor$estimate, cor$p.value)
  ) +
  
  theme(
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10),
    legend.key.width = unit(0.01, 'npc'),
    legend.key.height = unit(0.02, 'npc'),
  )

```

# 6. Plasma cytokines *in vitro*
```{r}
#| fig-width: 5
#| fig-height: 4

source('reviews/ggheatmap_helpers.R')

cyto <- c('IFNg', 'IL-10', 'G-CSF', 'IL-1a', 'TNFa', 'IL-1b', 'IL-1RA')
meta <- readr::read_rds('data/processed/invitro/metadata_cytokines_WBA.rds')
data <- readr::read_rds('data/processed/invitro/cytokines_of_interst_WBA.rds')

# make pretty labels
labs <- c(
  'IFNg'   = expression(IFN-gamma), 
  'IL-10'  = expression(IL-10), 
  'G-CSF'  = expression(G-CSF), 
  'IL-1a'  = expression(IL-1*alpha), 
  'TNFa'   = expression(TNF-alpha), 
  'IL-1b'  = expression(IL-1*beta), 
  'IL-1RA' = expression(IL-1*RA)
)

df <- data %>% 
  as_tibble(rownames = NA) %>% 
  rownames_to_column('DonorID') %>% 
  gather(var, val, -DonorID) %>% 
  filter(var %in% cyto) %>% 
  left_join(meta)

# paired t-test
stats <- df %>% 
  nest(data = -var) %>% 
  mutate(broom = data %>% map(~ lm(val ~ ID + Stimulation, data = .)) %>% map(broom::tidy)) %>% 
  unnest(broom) %>% 
  filter(!grepl('Intercept|ID', term)) %>% 
  transmute(var, Stimulation = gsub('Stimulation', '', term), pval = p.value) %>%
  annot_table('pval')

d <- df %>% 
  group_by(Stimulation, var) %>% summarise(val = mean(val), .groups = 'drop') %>% 
  mutate(val = scale(val)[ , 1], .by = var)

d <- d %>% 
  ggplot(aes(x = Stimulation, y = var)) +
  geom_tile(aes(fill = val), colour = 'white', size = 0.5) +
  geom_text(data = stats, aes(label = pval.signif), size = 4, vjust = 0.75, colour = 'white') +
  scale_fill_viridis_c(option = 'D', breaks = seq(-1, 1, by = 0.25), labels = c('-1.0', '', '-0.5', '', '0.0', '', '0.5', '', '1.0'), name = 'SLC') +
  coord_cartesian(expand = 0) +
  scale_x_discrete(position = 'top', labels = c('VEH', 'HBV', 'BCG', 'HBV+\nBCG')) +
  scale_y_discrete(position = 'left', labels = labs) +
  guides(fill = guide_colorsteps(nbin = 10)) +
  theme(
    legend.text = element_text(size = 9, hjust = 1),
    legend.title = element_text(size = 10),
    legend.key.width = unit(0.01, 'npc'),
    # legend.key.height = unit(0.02, 'npc'),
    legend.key.height = unit(0.05, 'npc'),
    axis.text = element_text(size = 9),
    axis.ticks = element_blank(), 
    axis.line = element_blank(), 
    axis.title = element_blank()
  )

d <- d %>% insert_hclust(cluster_by = val, side=c('right'), dist_method = 'euclidean', hclust_method = 'ward.D2')
  
cytokines <- grid::grid.draw(d)

```

# 7. Pathway PCAs *in vitro*
```{r}
#| column: page-right
#| fig-width: 12
#| fig-height: 4.5

e <- readr::read_rds('results/other_outputs/invitro_pathway_permanova.rds')

# filter and reorder
e <- e %>% filter(pathway_id %in% c('R-HSA-6798695', 'R-HSA-9020702', 'R-HSA-76002', 'R-HSA-6785807')) %>% slice(c(2, 4, 3, 1))

# get labels
e$pval <- e$pcaPlots %>% map(ggplot2::ggplot_build) %>% map('data') %>% map(3) %>% map('label') %>% map_dbl(unique)
e <- e %>% annot_table('pval')

# select just the plot data
e <- e %>% transmute(pathway_id, label, pval.signif, data = map(pcaPlots, 'data'))

# ugly hack to rotate PCs such that VEH is on the lower-left
# fix orientation manually (1)
e$data[[1]]$PC2 <- -e$data[[1]]$PC2
# fix orientation manually (2)
e$data[[2]]$PC1 <- -e$data[[2]]$PC1
# fix orientation manually (3) - no need
# fix orientation manually (4)
e$data[[4]]$PC1 <- -e$data[[4]]$PC1
e$data[[4]]$PC2 <- -e$data[[4]]$PC2

# re-order
e <- e %>% mutate(name = sprintf('%s\n%s', pathway_id, stringr::str_wrap(label, 25)))

pca <- e %>%
  pmap(\(name, data, pval.signif, ...) {
    data %>%
      mutate(stim = factor(Stim, levels = c('VEH', 'HBV', 'BCG', 'COMBO'), labels = c('VEH', 'HBV', 'BCG', 'HBV+\nBCG'))) %>% 
      ggplot(aes(PC1, PC2, colour = stim)) +
      geom_point(size = 2, alpha = 0.5) +
      ggpubr::stat_conf_ellipse(level = 0.95, linewidth = 0.75) +
      scale_color_manual(values = c('#888888', '#3C5488FF', '#E64B35FF', '#6F286AFF')) +
      
      # annotate(geom = 'text', x = max(data$PC1), y = max(data$PC2), hjust = 1, vjust = 1, size = 3.5, label = sprintf('p<%.2g', pval.signif)) +
      annotate(geom = 'text', x = max(data$PC1), y = max(data$PC2), hjust = 1, vjust = 1, size = 5, label = pval.signif) +
      labs(subtitle = name) +
    
      theme(
        legend.text = element_text(size = 9),
        legend.title = element_text(size = 10),
        legend.key.width = unit(0.01, 'npc'),
        legend.key.height = unit(0.02, 'npc'),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank()
      )
    
  }) %>% patchwork::wrap_plots(nrow = 1) + patchwork::plot_layout(guides = 'collect')

pca

```

# 8. Natural Experiment (*in vivo*) Results
```{r nat-expt-dev}

targets_invivo <- c(targets, 'MA-DOL30')
nat_expt_pal <- c("#440154FF", "#FDE725FF", "#3B528BFF", "#21908CFF", "#5DC863FF")

tmp <- con %>% 
  filter(query == 'anyhbv0_bcg7' | query == 'MA-DOL30') %>% 
  select(site, query, gsea, ssgsea) %>% 
  arrange(query, site)

tmp_a <- tmp %>% 
  select(site, query, gsea) %>% 
  unnest(gsea) %>% 
  filter(pathway %in% targets_invivo) %>%
  nest(gsea = -pathway)
  
tmp_b <- tmp %>% 
  filter(query != 'MA-DOL30') %>% 
  select(site, query, ssgsea) %>% 
  mutate(ssgsea = ssgsea %>% map(mutate, dol = as.character(dol))) %>% 
  unnest(ssgsea) %>% 
  filter(pathway %in% targets_invivo) %>%
  nest(ssgsea = -pathway)

nat_expt <- full_join(tmp_a, tmp_b) %>% arrange(match(pathway, targets_invivo))
rm(tmp, tmp_a, tmp_b)

# rename
nat_expt$pathway <- tibble(pathway = nat_expt$pathway) %>% 
  left_join(select(msigdb, pathway = description, pathway_id = geneset)) %>%
  mutate(pathway_id = ifelse(pathway_id == pathway, 'Module:', pathway_id)) %>% 
  mutate(label = sprintf('%s\n%s', pathway_id, stringr::str_wrap(pathway, 25))) %>%
  pull(label)

nat_expt$pathway[nat_expt$pathway == 'Module:\nMA-DOL30'] <- 'HARP\nModule'

nat_expt_1 <- nat_expt %>%
  pmap(\(pathway, gsea, ssgsea) {

    annot <- gsea %>%
      mutate(padj = p.adjust(pval, method = 'BH')) %>%
      annot_table('padj') %>%
      pmap(\(query, NES, padj, padj.signif, ...) ifelse(query == 'MA-DOL30', padj.signif, sprintf('NES\n%.2f\n%s', NES, padj.signif))) %>%
      unlist()

    # gsea DOL30
    l <- gsea %>%
      filter(query == 'MA-DOL30') %>%
      ggplot(aes(NES, query)) +
      geom_col(colour = nat_expt_pal[1], fill = nat_expt_pal[1], alpha = 0.5, width = 0.5) +
      geom_vline(xintercept = 0) +

      scale_x_continuous(limits = c(-3, 3), expand = expansion(mult = c(0.2, 0.05)), position = 'top') +
      theme(axis.title.y = element_blank(), axis.line.y = element_blank(), axis.ticks.y = element_blank()) +

      ggsignif::geom_signif(
        size = 0.3,
        textsize = 4,
        lineheight = 0.75,
        colour = 'black',
        xmin       = c(1),
        xmax       = c(1),
        y_position = gsea$NES[1],
        tip_length = 0,
        vjust = 2,
        annotation = annot[1],
      )

    # ssgsea DOL7 vs. DOL0
    r <- ssgsea %>%
      mutate(
        NES = scales::rescale(nes),
        dol = factor(dol, levels = c('0', '7'), labels = c('HBV', 'BCG'))
      ) %>%
      ggplot(aes(site, NES)) +
      geom_point(aes(colour = dol, group = dol), position = position_jitterdodge(jitter.width = 0.25, dodge.width = 0.5), pch = 21, alpha = 0.5, size = 0.5) +
      stat_summary(fun.data = 'mean_cl_boot', aes(colour = dol, fill = dol, group = dol), geom = 'pointrange', size = 0.25, position = position_dodge(width = 1.5), show.legend = F) +

      scale_fill_discrete (name = 'arm', palette = nat_expt_pal[c(3, 5)]) +
      scale_color_discrete(name = 'arm', palette = nat_expt_pal[c(3, 5)]) +
      scale_y_continuous(expand = expansion(mult = c(0.05, 0.25))) +

      labs(subtitle = pathway) +
      guides(color = guide_legend(override.aes = list(pch = 20, size = 3, alpha = 1))) +

      ggsignif::geom_signif(
        size = 0.3,
        textsize = 4,
        colour = 'black',
        xmin       = c(0.85, 1.85),
        xmax       = c(1.15, 2.15),
        y_position = 1.05,
        tip_length = c(0.01, 0.10, 0.01, 0.10),
        annotation = annot[-1],
      )

    if(!grepl('R-HSA', pathway)) {
      # patchw <- patchwork::wrap_plots(l + labs(subtitle = pathway), r + labs(subtitle = NULL), ncol = 1, heights = c(1, 5))
      l_tight <- l + 
        labs(subtitle = "HARP Module") +
        theme(
          plot.subtitle = element_text(
            hjust = 0, 
            vjust = 1,
            margin = margin(b = -35, unit = "pt") # Stronger negative pull to lift the bar plot
          ),
          plot.margin = margin(t = 2, r = 0, b = 0, l = 0, unit = "pt")
        )
      
      patchw <- patchwork::wrap_plots(
        l_tight, 
        r + labs(subtitle = NULL), 
        ncol = 1, 
        heights = c(0.5, 5) 
      )
    } else {
      patchw <-  r
    }

    patchw &
      labs(x = '') &
      theme(
        plot.margin = unit(c(0,0,0,0), 'npc'),
        axis.text = element_text(size = 9),
        legend.text = element_text(size = 9),
        legend.title = element_text(size = 10),
        legend.key.width = unit(0.01, 'npc'),
        legend.key.height = unit(0.02, 'npc'),
      )
  })

```

```{r nat-expt-tbl, eval=FALSE}

tmp <- con %>% 
  filter(query == 'anyhbv0_bcg7') %>%
  mutate(gsea = set_names(gsea, site)) %>% 
  pull(gsea) %>%
  map(export_gsea, msigdb)

tmp %>% 
  iwalk(~ {
    
    filepath <- 'results/tables'
    filename <- sprintf('natural_expt_%s_fgsea_results.csv', tolower(.y))
    
    readr::write_csv(
      .x,
      file = file.path(filepath, filename)
    )

    aws.s3::put_object(
      file.path(filepath, filename),
      object = file.path('s3://epichipc-main-dataintegration/Casey', filename)
    )
    
  })

```

```{r nat-expt-titers}

# just grab titer data we already collected for ssGSEA analysis...
titers_sub <- nat_expt$ssgsea[[5]] %>% 
  select(site, uid, sex, grp, grp_day = daygrp, DOL30, status) %>%
  mutate(DOL30 = exp(DOL30) + 0.25, grp = factor(grp, levels = c('HBV', 'HBV+BCG', 'BCG'))) %>% 
  filter(
    site == 'GAM',
    DOL30 <= 1e4 # instrument ULOQ
  )

library("scales")
reverselog_trans <- function(base = exp(1)) {
    trans <- function(x) -log(x, base)
    inv <- function(x) base^(-x)
    trans_new(paste0("reverselog-", format(base)), trans, inv, log_breaks(base = base), domain = c(1e-100, Inf))
}

nat_expt_2 <- titers_sub %>% 
  ggplot(aes(grp, DOL30, group = grp, colour = grp)) +
  geom_point(aes(shape = site), position = position_jitter(width = 0.15), size = 0.5, show.legend = F) +
  stat_summary(fun = 'mean', geom = 'col', aes(fill = grp), width = 0.5, alpha = 0.5) +
  stat_summary(fun.data = 'mean_cl_boot', stroke = 1, size = 0.5) +
  geom_hline(yintercept = 1) +
  geom_hline(yintercept = 10, linetype = 'dotted') +

  ggsignif::geom_signif(
    colour = 'black',
    size = 0.3,
    textsize = 4,
    
    vjust = -0.5,
    y_position = 2.25,
    tip_length = c(0.025, 0.25),
    
    comparisons = list(c('BCG', 'HBV')), 
    map_signif_level = function(x) paste0("p=", scales::pvalue(x)), test.args = list(exact = F)
  ) +

  scale_fill_manual  (values = nat_expt_pal[c(3, 5)]) +
  scale_colour_manual(values = nat_expt_pal[c(3, 5)]) +

  scale_x_discrete(labels = c('HBV', 'BCG')) +
  scale_y_continuous(
    breaks = c(1, 10, 100, 1000), 
    trans = 'log10',
    expand = expansion(mult = c(0.05, 0.10))
  ) +

  scale_shape_manual(values = c(21, 22)) +
  # facet_wrap(~ grp_day, nrow = 1) +
  labs(subtitle = 'Response\nDOL30', x = '', y = 'HBsAb titer (mIU/mL)') +
  theme(legend.position = 'none')

```

```{r other-tbls, eval=FALSE}

gsea <- readr::read_rds('results/other_outputs/gsea_tables.rds')

tmp <- gsea %>% 
  filter(grepl('^MA', query), site == 'GAM') %>% 
  pull(gsea) %>% 
  map(export_gsea, msigdb)

names(tmp) <- names(tmp) %>% gsub('-', 'minus_', .) %>% gsub('\\+', 'plus_', .)

tmp %>% 
  iwalk(~ {
    
    filepath <- 'results/tables'
    filename <- sprintf('baseline_%s_fgsea_results.csv', tolower(.y))
    
    readr::write_csv(
      .x,
      file = file.path(filepath, filename)
    )

    aws.s3::put_object(
      file.path(filepath, filename),
      object = file.path('s3://epichipc-main-dataintegration/Casey', filename)
    )
    
  })

```

# 9. Figure

### Top
```{r}
#| fig-height: 5
#| fig-width:  13
#| column: page-right

library(patchwork)

wrap_plots(
  plots_abstatus$invivo, 
  plots_abstatus$invitro, 
  free(plots_abstatus$common),
  nrow = 1, 
  widths = c(2, 2, 3)
) + plot_annotation(tag_levels = list(c('a', 'b', 'c')))

```

### Middle
```{r}
#| fig-height: 4
#| fig-width:  13
#| column: page-right

library(patchwork)

wrap_plots(free(pca), cytokines, widths = c(5, 1)) + 
  plot_layout(guides = 'collect') + 
  plot_annotation(tag_levels = list(c('d', '', '', '', 'e')))

```

### Bottom
```{r}
#| fig-height: 5.5
#| fig-width:  13
#| column: page-right

library(patchwork)

nat_expt_1 <- map(nat_expt_1, ~ . & theme(axis.text = element_text(size = 8)))
nat_expt_2 <- nat_expt_2 + theme(axis.text = element_text(size = 8))

bot <- nat_expt_1[1:4] %>% wrap_plots(nrow = 1)
bot <- wrap_plots(nat_expt_2, bot, last(nat_expt_1), widths = c(1, 5, 1.5), nrow = 1)

wrap_plots(bot) +
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = list(c('f', 'g', '', '', '', 'h', '', '')))

```

# 10. Mediation-analysis

```{r mediation-analysis-dol30}

# 0. Maternal Factors & Mediator
# ---------------------------------------------------------
maternal <- bind_rows(
  readr::read_rds('data/processed/gam/import_maternal.rds') %>% select(subject_id, birth_weight, maternal_age),
  readr::read_rds('data/processed/png/import_maternal.rds') %>% select(subject_id, birth_weight, maternal_age)
)

med <- nat_expt %>% select(pathway, ssgsea) %>% deframe() %>% last()

# 1. Preprocessing & Factoring
# ---------------------------------------------------------
dat <- med %>% 
  transmute(
    subject_id = uid, 
    sex = factor(sex),
    site = site,
    daygrp = ifelse(is.na(daygrp), 'DOL7', daygrp),
    treatment_group = factor(grp, levels = c('HBV', 'HBV+BCG', 'BCG'), labels = c('HBV', 'HBV', 'BCG')),
    
    pathway_score = nes,
    pathway_std    = as.numeric(scale(pathway_score)),
    
    outcome = DOL30,
    outcome_std    = as.numeric(scale(outcome)),
    outcome_cat    = exp(outcome) >= 10
  ) %>% 
  inner_join(maternal) %>%
  drop_na()

dat <- dat %>% 
  select(y = outcome_cat, mediator = pathway_std, pathway_score, trt = treatment_group, sex, birth_weight, maternal_age, site, daygrp) %>%
  mutate(sex = forcats::fct_rev(sex))

formula_1  <- y        ~                  trt + sex + maternal_age + birth_weight + site + daygrp
formula_2  <- y        ~       mediator + trt + sex + maternal_age + birth_weight + site + daygrp
formula_3  <- mediator ~                  trt + sex + maternal_age + birth_weight + site + daygrp
formula_2a <- y        ~ sex * mediator + trt * sex + maternal_age + birth_weight + site + daygrp
formula_3a <- mediator ~                  trt * sex + maternal_age + birth_weight + site + daygrp


# 2. Fit Standardized Models
# ---------------------------------------------------------
# Path a: Treatment -> Mediator | Path b & c': Mediator + Treatment -> Outcome
out_mod <- glm(formula_2, data = dat, family = binomial(link = "logit"))
med_mod <- lm(formula_3, data = dat)

set.seed(999)
fit_med <- mediation::mediate(
  med_mod, 
  out_mod, 
  treat = "trt", 
  mediator = "mediator",
  control.value = "HBV", 
  treat.value = "BCG", 
  sims = 1000, 
)


# 3. What about sex?
# ---------------------------------------------------------
out_mod_int <- glm(formula_2a, data = dat, family = binomial(link = "logit"))
med_mod_int <- lm(formula_3a, data = dat)

# 2. Estimate mediation for FEMALES
set.seed(999)
fit_female <- mediation::mediate(
  med_mod_int, 
  out_mod_int, 
  treat = "trt", 
  mediator = "mediator",
  control.value = "HBV", 
  treat.value = "BCG", 
  covariates = list(sex = "Female"),
  sims = 1000
)


# 3. Estimate mediation for MALES
set.seed(999)
fit_male <- mediation::mediate(
  med_mod_int, 
  out_mod_int, 
  treat = "trt", 
  mediator = "mediator",
  control.value = "HBV", 
  treat.value = "BCG", 
  covariates = list(sex = "Male"),
  sims = 1000
)


```

```{r mediation-analysis-dol128, include=FALSE, eval=FALSE}

# 0. Maternal Factors & Mediator
# ---------------------------------------------------------
maternal <- bind_rows(
  readr::read_rds('data/processed/gam/import_maternal.rds') %>% select(subject_id, birth_weight, maternal_age),
  readr::read_rds('data/processed/png/import_maternal.rds') %>% select(subject_id, birth_weight, maternal_age)
)

med <- nat_expt %>% select(pathway, ssgsea) %>% deframe() %>% last()

# 1. Preprocessing & Factoring
# ---------------------------------------------------------
dat <- med %>% 
  transmute(
    subject_id = uid, 
    sex = factor(sex),
    site = site,
    daygrp = ifelse(is.na(daygrp), 'DOL7', daygrp),
    treatment_group = factor(grp, levels = c('HBV', 'HBV+BCG', 'BCG'), labels = c('HBV', 'HBV', 'BCG')),
    
    pathway_score = nes,
    pathway_std    = as.numeric(scale(pathway_score)),
    
    outcome = DOL128,
    outcome_std    = as.numeric(scale(outcome)),
    outcome_cat    = exp(outcome) >= 10
  ) %>% 
  inner_join(maternal) %>%
  drop_na()

dat <- dat %>% 
  select(y = outcome_std, mediator = pathway_std, trt = treatment_group, sex, birth_weight, maternal_age, site, daygrp) %>%
  mutate(sex = forcats::fct_rev(sex))

formula_1  <- y        ~                  trt + sex + maternal_age + birth_weight + site + daygrp
formula_2  <- y        ~       mediator + trt + sex + maternal_age + birth_weight + site + daygrp
formula_3  <- mediator ~                  trt + sex + maternal_age + birth_weight + site + daygrp
formula_2a <- y        ~ sex * mediator + trt * sex + maternal_age + birth_weight + site + daygrp
formula_3a <- mediator ~                  trt * sex + maternal_age + birth_weight + site + daygrp


# 2. Fit Standardized Models
# ---------------------------------------------------------
# Path a: Treatment -> Mediator | Path b & c': Mediator + Treatment -> Outcome
out_mod <- lm(formula_2, data = dat)
med_mod <- lm(formula_3, data = dat)

set.seed(999)
fit_med <- mediation::mediate(
  med_mod, 
  out_mod, 
  treat = "trt", 
  mediator = "mediator",
  control.value = "HBV", 
  treat.value = "BCG", 
  sims = 1000, 
)


# 3. What about sex?
# ---------------------------------------------------------
out_mod_int <- lm(formula_2a, data = dat)
med_mod_int <- lm(formula_3a, data = dat)

# 2. Estimate mediation for FEMALES
set.seed(123)
fit_female <- mediation::mediate(
  med_mod_int, 
  out_mod_int, 
  treat = "trt", 
  mediator = "mediator",
  control.value = "HBV", 
  treat.value = "BCG", 
  covariates = list(sex = "Female"),
  sims = 1000
)


# 3. Estimate mediation for MALES
set.seed(123)
fit_male <- mediation::mediate(
  med_mod_int, 
  out_mod_int, 
  treat = "trt", 
  mediator = "mediator",
  control.value = "HBV", 
  treat.value = "BCG", 
  covariates = list(sex = "Male"),
  sims = 1000
)


```

::: {.callout-tip collapse="true"}
## Models

::: {.panel-tabset}
### Overall
```{r}
print(med_mod)
print(out_mod)
summary(fit_med)

```
### Females
```{r}
print(med_mod_int)
print(out_mod_int)
summary(fit_female)

```
### Males
```{r}
print(med_mod_int)
print(out_mod_int)
summary(fit_male)

```
:::

:::

```{r}

# reset
theme_set(
  cowplot::theme_cowplot() + 
    theme(
      strip.background = element_blank(), 
      axis.title = element_text(size = 12), 
      axis.text = element_text(size = 11)
    )
)

```

```{r plot-mediation}

# 1. Consolidate ONLY Mediation Results for Forest Plot
# ------------------------------------------------------------------
plot_data <- bind_rows(
  # --- OVERALL MODEL ---
  data.frame(
    term      = c("Mediation\n(ACME)", "Direct\n(ADE)", "Total Effect"),
    estimate  = c(fit_med$d0, fit_med$z0, fit_med$tau.coef),
    low       = c(fit_med$d0.ci[1], fit_med$z0.ci[1], fit_med$tau.ci[1]),
    high      = c(fit_med$d0.ci[2], fit_med$z0.ci[2], fit_med$tau.ci[2]),
    p_val     = c(fit_med$d0.p, fit_med$z0.p, fit_med$tau.p),
    Group     = "Overall"
  ),
  
  # --- FEMALE STRATUM ---
  data.frame(
    term      = c("Mediation\n(ACME)", "Direct\n(ADE)", "Total Effect"),
    estimate  = c(fit_female$d0, fit_female$z0, fit_female$tau.coef),
    low       = c(fit_female$d0.ci[1], fit_female$z0.ci[1], fit_female$tau.ci[1]),
    high      = c(fit_female$d0.ci[2], fit_female$z0.ci[2], fit_female$tau.ci[2]),
    p_val     = c(fit_female$d0.p, fit_female$z0.p, fit_female$tau.p),
    Group     = "Female"
  ),
  
  # --- MALE STRATUM ---
  data.frame(
    term      = c("Mediation\n(ACME)", "Direct\n(ADE)", "Total Effect"),
    estimate  = c(fit_male$d0, fit_male$z0, fit_male$tau.coef),
    low       = c(fit_male$d0.ci[1], fit_male$z0.ci[1], fit_male$tau.ci[1]),
    high      = c(fit_male$d0.ci[2], fit_male$z0.ci[2], fit_male$tau.ci[2]),
    p_val     = c(fit_male$d0.p, fit_male$z0.p, fit_male$tau.p),
    Group     = "Male"
  )
) %>%
  mutate(
    # Clean order: Total Effect at the top, then Direct, then Mediation at the bottom
    term = factor(term, levels = c("Mediation\n(ACME)", "Direct\n(ADE)", "Total Effect")),
    Group = factor(Group, levels = c("Overall", "Female", "Male")),
    p_label = ifelse(p_val < 0.001, "p<0.001", paste0("p=", round(p_val, 3)))
  )

# 2. Clean Visualization Focus Only on Mediation
# ------------------------------------------------------------------
pd <- position_dodge(width = 0.6)
annot <- count(dat, site)
annot <- sprintf('%s n=%d, %s n=%d', annot$site[1], annot$n[1], annot$site[2], annot$n[2])

p2_mediation <- ggplot(plot_data, aes(x = term, y = estimate, ymin = low, ymax = high, color = Group, fill = Group)) +
  geom_hline(yintercept = 0, linetype = "dotted", alpha = 0.5) +
  geom_linerange(size = 0.5, aes(group = Group), position = pd, show.legend = F) + 
  geom_point(colour = 'black', pch = 21, size = 3, position = pd) +
  coord_flip() +

  # Colored p-values shifting perfectly with the points
  geom_text(
    aes(label = p_label, y = high, color = Group), 
    position = pd, 
    hjust = -0.15, 
    size = 3.8, 
    show.legend = FALSE
  ) +
  
  # Adjust padding Dynamically based on the actual max data point
  expand_limits(y = max(plot_data$high) * 1.4) +
  
  scale_fill_manual(values = c("Overall" = "gray40", "Female" = "#e41a1c", "Male" = "#377eb8")) +
  scale_color_manual(values = c("Overall" = "gray40", "Female" = "#e41a1c", "Male" = "#377eb8")) +
  
  labs(
    title = "Vaccine Group → HARP Module → HBsAb DOL30", 
    subtitle = sprintf("Mediation Effects | Sex; %s", annot),
    x = NULL, 
    y = "Standardized Estimate (β)", 
    color = "Subgroup"
  ) +
  # theme_minimal(base_size = 13) + 
  theme(
    axis.line.y = element_blank(), 
    axis.ticks.y = element_blank(),
    legend.position = "top",
    plot.title = element_text(face = "plain", size = 14), # Matches Panel a title size
    plot.subtitle = element_text(size = 11, colour = "gray30")
  )

```

```{r plot-mediator}

library(lme4)
library(lmerTest)
library(emmeans)
library(tidyverse)
library(patchwork)
library(ggpubr)

# fit linear model
fit_b <- lm(pathway_score ~ sex * trt, data = dat)

# extract exact contrasts matching your 4 comparisons
lm_b_contrasts <- bind_rows(
  as_tibble(emmeans(fit_b, pairwise ~ trt | sex, adjust = 'fdr')$contrasts),
  as_tibble(emmeans(fit_b, pairwise ~ sex | trt, adjust = 'fdr')$contrasts)
) %>%
  mutate(
    group1     = c("Male\nHBV", "Female\nHBV", "Female\nHBV", "Female\nBCG"),
    group2     = c("Male\nBCG", "Female\nBCG", "Male\nHBV",   "Male\nBCG"),
    y.position = c(1.10,        1.03,          1.38,          1.24),
    p.format   = sprintf("p = %.2g", p.value)
  ) %>%
  select(group1, group2, p.format, y.position)

# ==============================================================================
# panel b plot generation
# ==============================================================================
p_mediator_final <- dat %>%
  mutate(
    sex = forcats::fct_rev(sex),
    group_label = paste(sex, trt),
    group_label = factor(group_label, 
                         levels = c("Female HBV", "Female BCG", "Male HBV", "Male BCG"), 
                         labels = c("Female\nHBV", "Female\nBCG", "Male\nHBV", "Male\nBCG")),
    trt = factor(trt, levels = c("HBV", "BCG"))
  ) %>%
  
  ggplot(aes(x = group_label, y = scales::rescale(mediator), alpha = trt, fill = sex)) + 
  ggrain::geom_rain(point.args = list(pch = 21, show.legend = FALSE), boxplot.args = list(show.legend = FALSE)) +
  
  # map the rank model results via explicit brackets
  ggpubr::stat_pvalue_manual(
    lm_b_contrasts,
    label = "p.format",
    size = 4,
    bracket.size = 0.5,
    tip.length = 0.02
  ) +
  
  scale_alpha_manual(values = c("HBV" = 0.5, "BCG" = 0.1), guide = 'none') +
  scale_color_manual(values = c("Female" = "#e41a1c", "Male" = "#377eb8")) +
  scale_fill_manual(values = c("Female" = "#e41a1c", "Male" = "#377eb8")) +

  labs(
    title = "HARP Module (Mediator)", 
    subtitle = "Baseline Expression | Sex,Vaccine Group",
    x = NULL, 
    y = "HARP Module", 
    fill = "Sex"
  ) +
  
  guides(colour = 'none') +
  theme(
    legend.position = "top",
    plot.title = element_text(face = "plain", size = 14), 
    plot.subtitle = element_text(size = 11, colour = "gray30")
  )

# overwrite alpha
p_mediator_final@layers[[4]]$aes_params$alpha <- 1

```

```{r, eval=FALSE, include=FALSE}
#| fig-height: 6
#| fig-width:  12
#| column: page-right

library(patchwork)

wrap_plots(p2_mediation, p_mediator_final) + plot_annotation(tag_level = 'a')
```

# 11. *In Vitro* Perturbations

```{r}
# File 1: Antibody concentration data (For Panels A and C)
file_path_ab <- "/data/epichipc-main/Clean_Data/In_vitro/MEMPHIS/Analysis_MultiOmic/HIPC HBV MEMPHIS ELISA antibody data_final analysis version_ST_10JUL2026.xlsx"
df_raw_ab <- readxl::read_excel(file_path_ab)

# File 2: IL-13 cytokine data (For Panel B)
file_path_cyto <- "/data/epichipc-main/Clean_Data/In_vitro/MEMPHIS/Analysis_MultiOmic/HIPC HBV MEMPHIS cytokine luminex data_final analysis version_ST_10JUL2026.xlsx"
df_raw_cyto <- readxl::read_excel(file_path_cyto, col_types = c(rep('guess', 5), 'numeric', 'numeric'))

# Clean-up
df_clean_ab <- df_raw_ab %>%
  transmute(
    PID = PID,
    Stimulation = relevel(factor(Stimulation), ref = 'PBS'),
    Inhibitor = recode(Inhibitor, "1:10 GATA3 inhb" = "GATA3"),
    Inhibitor = relevel(factor(Inhibitor), ref = 'no inhb'),
    Concentration
  )

df_clean_cyto <- df_raw_cyto %>%
  transmute(
    PID = PID,
    Stimulation = relevel(factor(Stimulation), ref = 'PBS'),
    Inhibitor = recode(Inhibitor, "GATA3 inhb 1:10" = "GATA3"),
    Inhibitor = relevel(factor(Inhibitor), ref = 'no inhb'),
    Concentration = IL_13
  )

```

<!-- ## Original Version -->
```{r, eval=FALSE, include=FALSE}
#| fig-height: 12
#| fig-width:  12
#| column: page-right

library(ggpubr)
library(rstatix)

pos <- position_jitter(width = 0.15, seed = 123)

# panel HBsAb induction
a <- df_clean_ab %>% filter(Inhibitor == 'no inhb') %>% arrange(Stimulation)

a_anova <- anova_test(data = a, dv = Concentration, wid = PID, within = Stimulation)
a_pwc <- a %>%
  pairwise_t_test(Concentration ~ Stimulation, paired = TRUE, p.adjust.method = "none") %>%
  filter(!(group1 == "HBV 1:10" & group2 == "HBV 1:50")) %>% 
  add_xy_position(x = "Stimulation")

a_panel <- a %>% 
  ggplot(aes(Stimulation, 10^Concentration, fill = Stimulation)) + 
  geom_path(position = pos, aes(group = PID), linewidth = 0.1, colour = 'black') +
  geom_point(pch = 21, size = 3, alpha = 0.5, position = pos) + 
  stat_summary(fun = 'mean', geom = 'point', pch = '_', size = 10, colour = 'black') +
  ggrepel::geom_text_repel(data = filter(a, Stimulation == 'PBS'), aes(label = PID), colour = 'black') +
  stat_pvalue_manual(a_pwc, label = "p = {p}", y.position = 1.1, hide.ns = FALSE, tip.length = 0.01, label.size = 4, step.increase = 0.10) +
  scale_fill_manual(values = c('PBS' = 'gray30', "HBV 1:10" = "#1D3557", "HBV 1:50" = "#377eb8"), guide = 'none') +
  scale_y_log10(limits = c(0.3, 20), breaks = scales::log_breaks(base = 10), labels = scales::label_comma()) +
  theme(legend.position = 'bottom') +
  facet_wrap(~ sprintf("RM-ANOVA p = %.2g", a_anova$ANOVA$p)) +
  labs(x = '', y = 'HBsAb (mIU/mL)')

# panel IL-13 abrogation
b <- df_clean_cyto

b_wilcox <- b %>%
  group_by(Stimulation) %>%
  wilcox_test(Concentration ~ Inhibitor, paired = TRUE) %>%
  add_significance() %>%
  add_xy_position(x = "Inhibitor")

b_panel <- b %>%
  ggplot(aes(Inhibitor, 10^Concentration)) + 
  geom_path(position = pos, aes(group = paste(PID, Stimulation)), linewidth = 0.1, colour = 'black') +
  geom_point(aes(fill = Inhibitor), pch = 21, size = 3, alpha = 0.5, position = pos) + 
  stat_summary(fun = 'mean', geom = 'point', pch = '_', size = 10, colour = 'black') +
  ggrepel::geom_text_repel(data = filter(b, Inhibitor == 'no inhb'), aes(label = PID), colour = 'black') +
  stat_pvalue_manual(b_wilcox, label = "p = {p}", y.position = 10, tip.length = 0.02, label.size = 4) +
  scale_fill_manual(values = c("no inhb" = "#377eb8", "GATA3" = "#e41a1c"), guide = 'none') +
  scale_y_log10(breaks = scales::log_breaks(base = 10), labels = scales::label_log(base = 10)) +
  facet_wrap(~ Stimulation) +
  theme(legend.position = 'bottom') +
  labs(x = '', y = 'IL-13 concentration')

# panel HBsAb abrogation
c <- df_clean_ab %>% filter(Stimulation == 'HBV 1:50')

c_wilcox <- c %>%
  wilcox_test(Concentration ~ Inhibitor, paired = TRUE) %>%
  add_significance() %>%
  add_xy_position(x = "Inhibitor")

c_panel <- c %>% 
  ggplot(aes(Inhibitor, 10^Concentration, fill = Inhibitor)) + 
  geom_path(position = pos, aes(group = paste(PID, Stimulation)), linewidth = 0.1, colour = 'black') +
  geom_point(pch = 21, size = 3, alpha = 0.5, position = pos) + 
  stat_summary(fun = 'mean', geom = 'point', pch = '_', size = 10, colour = 'black') +
  ggrepel::geom_text_repel(aes(label = ifelse(Inhibitor == 'no inhb', PID, NA)), colour = 'black') +
  stat_pvalue_manual(c_wilcox, label = "p = {p}", y.position = 1.27, tip.length = 0.02, label.size = 4) +
  scale_fill_manual(values = c("no inhb" = "#377eb8", "GATA3" = "#e41a1c"), guide = 'none') +
  scale_y_log10(limits = c(0.3, 20), breaks = scales::log_breaks(base = 10), labels = scales::label_comma()) +
  facet_wrap(~ Stimulation) +
  theme(legend.position = 'bottom') +
  labs(x = '', y = 'HBsAb (mIU/mL)')

d <- "AABB
      CDDE"

wrap_plots(design = d, p2_mediation, p_mediator_final, a_panel, b_panel, c_panel) + plot_annotation(tag_level = 'a')

```

## My Version
```{r}
# ==============================================================================
# 1. global setup
# ==============================================================================
set.seed(124)

library(nlme)
library(emmeans)
library(tidyverse)
library(patchwork)
library(ggpubr) 

# shared aesthetics
pos <- position_jitter(width = 0.15, seed = 1)

# file paths
file_path_ab   <- '/data/epichipc-main/Clean_Data/In_vitro/MEMPHIS/Analysis_MultiOmic/HIPC HBV MEMPHIS ELISA antibody data_final analysis version_ST_17JUL2026.xlsx'
file_path_cyto <- '/data/epichipc-main/Clean_Data/In_vitro/MEMPHIS/Analysis_MultiOmic/HIPC HBV MEMPHIS cytokine luminex data_final analysis version_ST_17JUL2026.xlsx'

# ==============================================================================
# 2. cleaning pipelines
# ==============================================================================

# antibody data
df_clean_ab <- readxl::read_excel(file_path_ab) %>%
  transmute(
    PID, 
    Stimulation = fct_relevel(factor(Stimulation), 'PBS', 'HBV 1:50', 'HBV 1:10'),
    Inhibitor = fct_relevel(recode(Inhibitor, 'no inhb' = 'Ctrl', '1:10 GATA3 inhb' = 'GATA3\n(-)', 'Inflammasome inhb' = 'INFLAM\n(-)'), 'Ctrl'),
    Concentration
  )

# cytokine data
df_clean_cyto <- readxl::read_excel(file_path_cyto) %>%
  select(PID, Stimulation, Inhibitor, all_of(c('IL.13', 'IL.9', 'IL.1a', 'IL.1b', 'IL.6', 'GRO'))) %>%
  pivot_longer(cols = c(-PID, -Stimulation, -Inhibitor), names_to = 'var', values_to = 'Concentration') %>%
  mutate(
    Stimulation = fct_relevel(factor(Stimulation), 'PBS'),
    Inhibitor   = fct_relevel(recode(Inhibitor, 'no inhb' = 'Ctrl', 'GATA3 inhb 1:10' = 'GATA3\n(-)', 'Inflammasome inhb' = 'INFLAM\n(-)'), 'Ctrl'),
    var         = recode(var, 'IL.13' = 'IL-13', 'IL.9' = 'IL-9', 'IL.1a' = 'IL-1A', 'IL.1b' = 'IL-1B', 'IL.6' = 'IL-6', 'GRO' = 'CXCL1')
  )

# ==============================================================================
# 3. statistical & plotting engines
# ==============================================================================

# multiplex cytokine panels (d and f)
plot_cytokines <- function(data, targets, palette_colors) {
  df_filtered <- data %>% 
    filter(var %in% targets, Inhibitor %in% names(palette_colors)) %>% 
    droplevels()
  
  # model engine using nlme::lme
  lm_contrast <- df_filtered %>%
    nest(data = -var) %>%
    mutate(fit = map(data, ~ {
      .x %>%
        mutate(Ranks = rank(Concentration, ties.method = "average")) %>%
        nlme::lme(Ranks ~ Stimulation * Inhibitor, random = ~ 1 | PID, data = .) %>%
        emmeans::emmeans(pairwise ~ Inhibitor | Stimulation, adjust = 'none') %>%
        pluck("contrasts") %>% 
        as_tibble()
    })) %>%
    unnest(fit)
  
  # plot engine
  df_filtered %>%
    ggplot(aes(Inhibitor, Concentration)) +
    geom_path(position = pos, aes(group = PID), linewidth = 0.1) +
    geom_point(position = pos, aes(fill = Inhibitor), pch = 21, size = 3, alpha = 0.5) +
    stat_summary(fun.data = 'mean_se', size = 0.5, colour = 'black') +
    geom_text(aes(x = 1.5, y = Inf, label = sprintf('p = %.2g', p.value)), vjust = 1.2, size = 4, data = lm_contrast) +
    scale_y_continuous(expand = expansion(mult = c(0.1, 0.2))) +
    scale_fill_manual(values = palette_colors, guide = 'none') +
    labs(x = '', y = expression(paste("Log"[2], "(pg/mL)"))) +
    theme(strip.text = element_text(size = 8)) +
    facet_wrap(~ interaction(var, Stimulation, sep = " | "), nrow = 2, scales = 'free_y')
}

# single pathway panels (e and g)
plot_ab_pathways <- function(data, target_inhibitor, facet_label, fill_color) {
  df_filtered <- data %>%
    filter(Stimulation == 'HBV 1:50', Inhibitor %in% c('Ctrl', target_inhibitor)) %>%
    droplevels()
  
  # model engine using nlme::lme
  lm_contrast <- df_filtered %>%
    mutate(Ranks = rank(Concentration, ties.method = "average")) %>%
    nlme::lme(Ranks ~ Inhibitor, random = ~ 1 | PID, data = .) %>%
    emmeans::emmeans(pairwise ~ Inhibitor, adjust = 'tukey') %>%
    pluck("contrasts") %>% 
    as_tibble()

  # safe color assignments
  fill_palette <- c('Ctrl' = '#377eb8')
  fill_palette[target_inhibitor] <- fill_color

  # plot engine
  df_filtered %>%
    ggplot(aes(Inhibitor, 10^Concentration)) +
    geom_path(position = pos, aes(group = PID), linewidth = 0.1) +
    geom_point(position = pos, aes(fill = Inhibitor), pch = 21, size = 3, alpha = 0.5) +
    stat_summary(fun.data = 'mean_se', size = 0.5, colour = 'black') +
    geom_text(aes(x = 1.5, y = Inf, label = sprintf('p = %.2g', p.value)), vjust = 1.2, size = 4.5, data = lm_contrast) +
    ggrepel::geom_text_repel(data = filter(df_filtered, Inhibitor == 'Ctrl'), aes(label = PID), size = 4, colour = 'black') +
    scale_fill_manual(values = fill_palette, guide = 'none') +
    scale_y_log10(expand = expansion(mult = c(0.05, 0.1)), labels = label_number(accuracy = 0.1)) +
    facet_wrap(~ factor(Stimulation, levels = c('PBS', 'HBV 1:50')), labeller = as_labeller(c('HBV 1:50' = facet_label))) +
    labs(x = '', y = 'HBsAb (mIU/mL)') +
    theme(legend.position = 'bottom')
}

# ==============================================================================
# 4. generation & layout
# ==============================================================================

# panels c, d
panel_c <- df_clean_cyto %>% 
  plot_cytokines(
    targets = c('IL-1A', 'IL-1B', 'IL-6', 'CXCL1'), 
    palette_colors = c('Ctrl' = '#377eb8', 'INFLAM\n(-)' = '#e41a1c')
  )

panel_d <- df_clean_cyto %>% 
  plot_cytokines(
    targets = c('IL-13', 'IL-9'), 
    palette_colors = c('Ctrl' = '#377eb8', 'GATA3\n(-)' = '#e41a1c')
  )

# panel e - this one is a bit different because we have a 3-level factor instead of a 2x2 design
df_panel_e <- df_clean_ab %>%
  filter(Inhibitor == 'Ctrl') %>%
  mutate(Stimulation = fct_relevel(factor(Stimulation), 'PBS'), Ranks = rank(Concentration, ties.method = "average"))

lm_e_contrast <- df_panel_e %>% 
  nlme::lme(Ranks ~ Stimulation, random = ~ 1 | PID, data = .) %>%
  emmeans::emmeans(trt.vs.ctrl ~ Stimulation, ref = "PBS") %>%
  pluck("contrasts") %>% 
  as_tibble() %>%
  transmute(
    group1 = "PBS", 
    group2 = str_remove(contrast, " - PBS"), 
    y.position = c(1.2, 1.25), 
    p = round(p.value, 3)
  )

panel_e <- df_panel_e %>%
  arrange(Stimulation) %>%
  ggplot(aes(Stimulation, 10^Concentration, fill = Stimulation)) +
  geom_path(position = pos, aes(group = PID), linewidth = 0.1, colour = 'black') +
  geom_point(pch = 21, size = 3, alpha = 0.5, position = pos) +
  stat_summary(fun.data = 'mean_se', size = 0.5, colour = 'black') +
  stat_pvalue_manual(lm_e_contrast, label = 'p = {p}', hide.ns = FALSE, label.size = 4.5, tip.length = 0.02, step.increase = 0.10) +
  ggrepel::geom_text_repel(data = filter(df_panel_e, Stimulation == 'PBS'), aes(label = PID), size = 4, colour = 'black') +
  scale_fill_manual(values = c('PBS' = 'gray30', 'HBV 1:10' = '#1D355f', 'HBV 1:50' = '#377eb8'), guide = 'none') +
  scale_y_log10(expand = expansion(mult = c(0.05, 0.1)), labels = label_number(accuracy = 0.1)) +
  labs(x = '', y = 'HBsAb (mIU/mL)') +
  theme(legend.position = 'bottom')

# panel f, g
panel_f <- df_clean_ab %>% 
  plot_ab_pathways(
    target_inhibitor = 'INFLAM\n(-)', 
    facet_label = 'HBV 1:50 ± INFLAM(-)', fill_color = '#e41a1c'
  )

panel_g <- df_clean_ab %>% 
  plot_ab_pathways(
    target_inhibitor = 'GATA3\n(-)', 
    facet_label = 'HBV 1:50 ± GATA3(-)', fill_color = '#e41a1c'
  )

```

```{r}
#| fig-width:  13
#| fig-height: 14.18
#| column: page-right

# grid assembly
design <- "
AAB
CCD
EFG
"

wrap_plots(
  free(p2_mediation, side = 'l'), p_mediator_final, 
  panel_c, panel_d, panel_e, 
  panel_f, panel_g, 
  design = design
) + 
plot_annotation(tag_level = 'a')

# ==============================================================================
# print mean differences for antibody panels
# ==============================================================================

# panel d: gata3 difference
mean_diff_d <- df_clean_ab %>%
  filter(Stimulation == "HBV 1:50", Inhibitor %in% c("Ctrl", "GATA3\n(-)")) %>%
  group_by(Inhibitor) %>%
  summarise(mean_val = mean(10^Concentration, na.rm = TRUE)) %>%
  pivot_wider(names_from = Inhibitor, values_from = mean_val) %>%
  mutate(diff = Ctrl - `GATA3\n(-)`) %>%
  pull(diff)

# panel e: inflammasome difference
mean_diff_e <- df_clean_ab %>%
  filter(Stimulation == "HBV 1:50", Inhibitor %in% c("Ctrl", "INFLAM\n(-)")) %>%
  group_by(Inhibitor) %>%
  summarise(mean_val = mean(10^Concentration, na.rm = TRUE)) %>%
  pivot_wider(names_from = Inhibitor, values_from = mean_val) %>%
  mutate(diff = Ctrl - `INFLAM\n(-)`) %>%
  pull(diff)

# print to console
cat(sprintf("Panel D (GATA3) Mean Difference (Ctrl - Inh): %.2f mIU/mL\n", mean_diff_d))
cat(sprintf("Panel E (INFLAM) Mean Difference (Ctrl - Inh): %.2f mIU/mL\n", mean_diff_e))

```

# 12. Post-vaccine Supplemental Figure
```{r}
#| fig-width:  8.25
#| fig-height: 9
#| column: page-right

library(tidyverse)
library(patchwork)

# ==============================================================================
# 1. PREP DATA
# ==============================================================================
foo <- readr::read_rds('data/processed/gam/import_luminex_cytokines.rds') %>% 
  as_tibble(rownames = NA) %>% 
  rownames_to_column('vid') %>% 
  gather(var, val, -vid)

bar <- readr::read_rds('data/processed/gam/import_metadata.rds') %>% 
  select(uid = `Unique Identifier`, vid = `Visit Identifier`, grp = VaccineGrp, daygrp = DayGrp, dol = DOL, sex = Sex)

# Isolate strictly to IL-1a
foobar <- foo %>% right_join(bar, .) %>% filter(var == 'IL-1a')

# Calculate baseline-subtracted differences (DiD data)
foobar_deltas <- foobar %>% 
  select(-vid) %>% 
  group_by(uid, var) %>% 
  filter(n() > 1) %>% 
  arrange(dol) %>% 
  mutate(dol = c('pre', 'post')) %>% 
  ungroup() %>% 
  spread(dol, val) %>% 
  mutate(diff = post - pre)

rm(foo, bar)

# ==============================================================================
# 2. OMNIBUS ANOVA (Your original analysis confirming IL-1a is significant)
# ==============================================================================
# Run the omnibus test to check for main effects and interactions
il1a_model <- lm(diff ~ grp * daygrp, data = foobar_deltas)
model_anova <- anova(il1a_model) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = 'Effect') %>% 
  filter(Effect %in% c('grp', 'grp:daygrp')) %>% 
  tibble()

# cat('\n=== OMNIBUS ANOVA (Confirms Overall Significance) ===\n')
# print(model_anova)

# Define label used in panel_il1a_did subtitle
anova_label <- sprintf("Overall vaccine effect: p=%.2g", model_anova$`Pr(>F)`[1])

# ==============================================================================
# 3. POST-HOC CONTRASTS (Breaking down where the differences are)
# ==============================================================================

# Setup the estimated marginal means grouped by timepoint
emm_by_day <- emmeans::emmeans(il1a_model, ~ grp | daygrp)

# --- TEST: Each individual vaccine group vs. Delayed at every timepoint ---
individual_vs_delayed <- contrast(emm_by_day, method = 'trt.vs.ctrl', ref = 'Delayed', adjust = 'fdr') %>% 
  as.data.frame() %>% 
  tibble()

# cat('\n=== TEST: Individual Vaccine Groups vs. Delayed Per Timepoint ===\n')
# print(individual_vs_delayed %>% arrange(p.value))

## ==============================================================================
# 4. CLEAN NET DID PLOT (FACETED BY DOL)
# ==============================================================================

# 1. Compute net contrasts properly using emmeans
did_contrasts <- emmeans::contrast(emm_by_day, method = "trt.vs.ctrl", ref = "Delayed", adjust = "fdr") %>%
  as.data.frame() %>%
  tibble() %>%
  mutate(
    grp = contrast %>% gsub('^\\((.*)\\).*', '\\1', .) %>% gsub(" - Delayed", "", .) %>% factor(levels = c("BCG", "HBV", "HBV+BCG")),
    daygrp = factor(daygrp, levels = c("DOL1", "DOL3", "DOL7"))
  )

# 2. Assign exact colors for the 3 remaining groups
did_colors <- c(
  'BCG'     = '#E04D26',
  'HBV'     = '#3B528B',
  'HBV+BCG' = '#6A2C70'
)

# 3. Define annotation coordinates with diagonal offset to the right
stat_df <- did_contrasts %>%
  filter(daygrp == "DOL3", grp == "BCG") %>% 
  mutate(
    daygrp  = factor(daygrp, levels = c("DOL1", "DOL3", "DOL7")),
    grp     = factor(grp, levels = c("BCG", "HBV", "HBV+BCG")),
    label   = paste0('FDR=', scales::pvalue(p.value, add_p = F)),
    
    # Point on top of BCG error bar (BCG = 1 on the x-axis factor scale)
    x_start = 1,
    y_start = estimate + SE + 0.01,
    
    # Text position shifted right toward panel middle (e.g., x = 1.6) and higher
    x_end   = 1.2,
    y_end   = estimate + SE + 0.045
  )

# 4. Plot panel B
panel_il1a_did <- ggplot(did_contrasts, aes(x = grp, y = estimate, colour = grp)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  
  geom_pointrange(aes(ymin = estimate - SE, ymax = estimate + SE), size = 0.6) +
  facet_wrap(~daygrp, nrow = 1) +
  scale_colour_manual(values = did_colors) +
  
  # --- DIAGONAL LINKING LINE (SLOPING RIGHT) ---
  geom_segment(
    data = stat_df,
    aes(x = x_start, xend = x_end - 0.025, y = y_start, yend = y_end - 0.01),
    inherit.aes = FALSE,
    colour = "black",
    linewidth = 0.35
  ) +
  
  # --- P-VALUE TEXT SHIFTED RIGHT ---
  geom_text(
    data = stat_df,
    aes(x = x_end, y = y_end, label = label),
    inherit.aes = FALSE,
    colour = "black",
    hjust = 0,     # Left-aligned so text flows rightward from x_end
    vjust = 0.2,   # Aligns vertically with the top end of the pointer line
    size = 3
  ) +
  
  coord_cartesian(ylim = c(-0.1, 0.26)) +
  cowplot::theme_cowplot() +
  theme(
    legend.position = 'bottom',
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    strip.background = element_blank(),
    panel.spacing = unit(1, "lines")
  ) +
  labs(
    subtitle = sprintf("IL-1α plasma kinetics\n%s", anova_label),
    x = NULL,
    y = expression(paste("Log"[2], " FC vs. Delayed")),
    colour = 'Vaccine group',
    # caption = "* BCG vs. Delayed FDR-adjusted"
  )

# =========================================================================
# 5. Load Extracted Data Frames
# =========================================================================
df          <- readr::read_tsv("data/processed/extracted_es_data.csv")
grp_fdr     <- readr::read_tsv("data/processed/extracted_grp_fdr.csv")
overall_fdr <- readr::read_tsv("data/processed/extracted_overall_fdr.csv")

# =========================================================================
# 6. Factor Level Ordering
# =========================================================================
pathway_levels         <- c("R-HSA-6798695", "R-HSA-9020702", "R-HSA-76002")
df$pathway_id          <- factor(df$pathway_id, levels = pathway_levels)
grp_fdr$pathway_id     <- factor(grp_fdr$pathway_id, levels = pathway_levels)
overall_fdr$pathway_id <- factor(overall_fdr$pathway_id, levels = pathway_levels)

time_levels        <- c("DOL1", "DOL3", "DOL7")
df$daygrp          <- factor(df$daygrp, levels = time_levels)
grp_fdr$daygrp     <- factor(grp_fdr$daygrp, levels = time_levels)
overall_fdr$daygrp <- factor(overall_fdr$daygrp, levels = time_levels)

grp_levels   <- c("HBV", "HBV+BCG", "BCG")
df$grp       <- factor(df$grp, levels = grp_levels)
grp_fdr$grp  <- factor(grp_fdr$grp, levels = grp_levels)

# =========================================================================
# 7. Plotting Loop with Hardcoded Text Coordinates
# =========================================================================
theme_set(cowplot::theme_cowplot() + theme(strip.background = element_blank(), legend.position = 'bottom'))

# df_downsampled <- df %>% group_by(pathway_id, daygrp) %>% sample_n(300) %>% ungroup() %>% arrange(pathway_id, daygrp, Rank)

df %>% 
  # df %>% 
  nest(data = -c(pathway_id, pathway_name)) %>% 
  pmap(\(pathway_id, pathway_name, data) {
    
    current_overall_fdr <- overall_fdr %>% filter(pathway_id == !!pathway_id)
    current_grp_fdr     <- grp_fdr     %>% filter(pathway_id == !!pathway_id)
    
    data %>% 
      # group_by(daygrp) %>% sample_n(250) %>% ungroup() %>% 
      arrange(Rank) %>% 
      ggplot(aes(Rank, ES, colour = grp)) +
      geom_hline(yintercept = 0, linetype = 'dotted') +
      geom_path(show.legend = F) +
      
      # Layer 1: Top p-value annotations (Hardcoded x = 50, fixed top y-axis)
      geom_text(
        data = current_overall_fdr,
        aes(x = 100, y = 0.55, label = p_label),
        color = "black",
        size = 3,
        hjust = 1,
        vjust = 0.5,
        inherit.aes = FALSE
      ) +
      
      # Layer 2: Bottom FDR symbols/stars (Stacked vertically with superscript 'ns')
      geom_text(
        data = current_grp_fdr,
        aes(
          x = 100, 
          y = 0.50 - (as.numeric(grp) * 0.075), 
          label = ifelse(symbol == "-", "'ns'^''", sprintf("'%s'", symbol)), 
          color = grp
        ),
        size = 3,
        hjust = 1,
        vjust = 0.5,
        parse = TRUE,        # Tells ggplot to interpret the text string as an expression
        show.legend = FALSE
      ) +
      
      scale_color_manual(values = c("HBV" = "#3c5488", "HBV+BCG" = "#6f286a", "BCG" = "#e64b35")) +
      scale_x_continuous(breaks = c(0, 50, 100), labels = function(x) paste0(x, "%")) +
      scale_y_continuous(expand = expansion(mult = c(0.1, 0.2))) + # Generates layout padding for text
      facet_wrap(~ daygrp, nrow = 1) +
      labs(subtitle = sprintf('%s\n%s', pathway_id, pathway_name), x = 'rank', y = 'ES')
  }) %>% 
  c(., panel_il1a_did) %>% 
  .[c(2, 4, 1, 3)] %>% 
  wrap_plots() + 
  plot_layout(guides = 'collect') +
  plot_annotation(tag_level = 'a') &
  theme(
    axis.text.x = element_text(size = 8), 
    plot.subtitle = element_text(size = 10),
    strip.text = element_text(size = 10),
    panel.spacing = unit(1, "lines"), 
  )

```

